### 2021-03-23 notes/draft_article.tex
.

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\graphicspath{{./figures/}}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=1.5cm, right=1.5cm, top=2.50cm, bottom=2.50cm]{geometry}
\usepackage{multicol}
\usepackage{float}

\author{Fabian Schubert}
\title{Nonlinear Dendritic Coincidence Detection for Supervised Learning}

\usepackage{natbib}
\bibliographystyle{apalike}

\renewcommand{\arraystretch}{1.5}

\usepackage{layouts}

\begin{document}
	\maketitle

		\section{Introduction}

		In recent years, a growing body of research has addressed the 
		functional implications of the distinct physiology and anatomy of 
		cortical pyramidal neurons. In particular, on the theoretical side,
		we saw a paradigm shift from treating neurons as point-like electrical
		structures towards embracing the entire dendritic structure. This was 
		mostly due to the fact that experimental work uncovered dynamical properties
		of these cells that simply could not be accounted for by point models.

		An important finding was that the apical dendritic tree of
		cortical pyramidal neurons can act as a separate nonlinear synaptic 
		integration zone. Under certain conditions, a dendritic $\rm Ca^{2+}$ spike
		can be elicited that propagates towards the soma, causing rapid, bursting
		spiking activity. One of the cases in which dendritic spiking can occur
		was termed 'backpropagation-activated $\rm Ca^{2+}$ spike firing' 
		('BAC firing'): A single somatic spike can backpropagate towards the apical
		spike initiation zone, in turn significantly facilitating the initiation of 
		a dendritic spike. This reciprocal coupling is believed to act as a form of
		coincidence detection: If apical and basal synaptic input co-occurs, the 
		neuron can respond with a rapid burst of spiking activity. The firing rate
		of these temporal bursts exceeds the firing rate that is maximally achievable 
		under basal synaptic input alone, therefore representing a form of temporal coincidence
		detection between apical and basal input.

		Naturally, these mechanisms also affect plasticity and thus learning
		within the cortex. While the interplay between basal and apical stimulation and
		its effect on synaptic efficacies is subject to ongoing research, there is
		some evidence that BAC-firing tends to shift plasticity towards long-term potentiation
		(LTP). Thus, coincidence between basal and apical input appears to also gate synaptic
		plasticity.

		In a supervised learning scheme, where the top down input
		arriving at the apical compartment acts as the teaching signal,
		the most straight-forward learning rule for the basal synaptic
		weights would be derived from an appropriate loss function,
		such as a mean square error, based on the difference between 
		basal and apical input, i.e. $I_p - I_d$. Theoretical work has 
		investigated possible learning mechanisms
		that could utilize an intracellular error signal
		\citep{Urbanczik_2014,Schiess_2016,Guerguiev_2017}.
		However, a clear experimental
		evidence for a physical quantity encoding such an error 
		is---to our knowledge---yet to be found. 
		On the other hand, Hebbian-type plasticity is extensively
		documented in experiments. Therefore, our work was
		based on the question whether the non-linear interactions
		between basal and apical synaptic input could, when
		combined with a Hebbian plasticity rule, allow a neuron
		to learn to reproduce an apical teaching signal in its
		proximal input.

		In our work, we combined a phenomenological model predicting the output
		firing rate as a function of two streams of synaptic input (subsuming basal and apical input)
		with a BCM-like plasticity rule on basal synapses. LTP was induced in our model 
		if the postsynaptic activity surpassed a threshold level that was only achievable if 
		both basal and apical inputs were present. Therefore, we hypothesized that 
		this combination of neural activation and plasticity rules would lead to an
		increased correlation between basal and apical input.

		Furthermore, this temporal alignment could potentially facilitate apical inputs to act
		as top-down teaching signals, without the need for an explicit error-driven
		learning rule. Thus, we also tested our model in a simple linear 
		supervised classification task and compared it to the performance of simple
		point neuron combined with a BCM-plasticity rule on the basal inputs.

		\section{Model}

		\subsection{Neuron Model}

		The neuron model used throughout this study is a discrete-time rate encoding model that
		uses two separate input variables, subsuming the total synaptic
		input current injected arriving at the basal (proximal) and apical (distal) 
		dendritic structure of a pyramidal neuron, respectively. The model is a slightly
		simplified version of a phenomenological model proposed by \citet{Shai_2015}.
		Denoting the input currents by $I_p$ (proximal) and $I_d$ (distal), 
		the model is written as
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
		\begin{split}
		y\left(t\right) &= \sigma\left( I_p(t) - \theta_{p0} \right)
		\left[1-\sigma\left(I_d(t) - \theta_d\right)\right] \\
		&+ \alpha \sigma\left(I_d(t) - \theta_d \right)\sigma\left( I_p(t) - \theta_{p1} \right)
		\end{split} \label{eq:comp_model}\\
		\sigma(x) &\equiv \frac{1}{1+\exp(-4x)} \; .
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $\theta_{p0}$, $\theta_{p1}$ and $\theta_d$ are threshold variables with
		respect to proximal and distal input. Overall, this equation describes two distinct
		regions of neural activation in the $(I_p, I_d)$-space which differ in their
		maximal firing rates, which are set to $1$ and $\alpha$, where $0 < \alpha < 1$.
		A plot of \eqref{eq:comp_model} is shown in Fig.~\ref{fig:comp_model}.
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\centering
			\includegraphics[width=0.8\columnwidth]{plot_comp_mod_marks.png}
			\caption{{\bf Two-compartment rate model.} Firing rate as a
			function of proximal and distal input $I_p$ and $I_d$, 
			see \eqref{eq:comp_model}. 
			The thresholds $\theta_{p0}$, $\theta_{p1}$ and $\theta_d$ define
			two regions of neural activity with a a maximal firing rate of $1$
			and $\alpha$.}
			\label{fig:comp_model}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		In all our numerical experiments, we compared this model with a 
		simple point neuron model that was given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			y(t) = \sigma\left(I_p(t) + I_d(t) - \theta \right) \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		In all our numerical experiments, the apical input $I_d$ was generated
		``as is", meaning, it was not dynamically calculated as a superposition
		of multiple presynaptic inputs, but given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			I_d(t) = n_d(t) x_d(t) - b_d(t) \; ,
			\label{eq:I_d}
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $n_d(t)$ is a scaling factor, $x_d(t)$ a pre-generated
		discrete time sequence and $b_d(t)$ a bias. Note that $n_d$ and $b_d$ 
		are time dependent since they were subject to adaptation processes 
		that are described in the next section.

		Similarly, $I_p(t)$ was given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			I_p(t) = n_p(t) \sum_{i=1}^{N} x_{p,i}(t) w_i(t) - b_p(t) \; ,
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $N$ is the number of presynaptic afferents, $x_{p,i}(t)$ the
		corresponding sequences and $w_i(t)$ the synaptic efficacies.
		As for $I_d(t)$, $n_p(t)$ and $b_p(t)$ was used as a time dependent
		scaling and bias.

		\subsection{Plasticity}

		We implemented a BCM-like plasticity rule for the basal 
		synaptic weights given by the following update equation:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			w_i(t+1) &= w_i(t) + \mu_w \left(x_{p,i}(t) - \tilde{x}_{p,i}(t)\right)
			y(t)\left[y(t) - \theta_y \right]
			\label{eq:bcm_plast} \\
			\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1)
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		While the classical BCM-rule dynamically modifies the 
		threshold $\theta_y$ to stabilize weights, we used a
		synaptic normalization constraint
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			w_i(t) \rightarrow \frac{w_i(t)}{\lVert \mathbf{w}(t)\rVert}
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		in each time step, where $\lVert \mathbf{w}(t)\rVert$ denotes the Euclidean
		norm of the synaptic weight vector. The threshold $\theta_y$ was
		chosen such that LTP is present if neural activity is in the
		high-activity regime (yellow area in Fig.~\ref{fig:comp_model}), while
		LTD is present in the low-activity regime 
		(given by $\alpha$, see Fig.~\ref{fig:comp_model}). Hence, we set
		$\theta_y = (1+\alpha)/2$.

		For comparative reasons, the point neuron model was equipped with the
		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}, 
		except that the threshold $\theta_y$ was dynamically adjusted as
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			\theta_y(t+1) = \left(1-\mu_{av}\right)\theta_y(t) + \mu_{av} y^2(t) \; , 
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		which is an implementation of a running average of the square activity,
		as used in the classic BCM-rule.

		Additionally, the scaling and bias variables were changing dynamically
		according to the following homeostatic plasticity rules:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			b_p(t+1) &= b_p(t) + \mu_b \left[I_p(t) - I_p^t\right] \\
			b_d(t+1) &= b_d(t) + \mu_b \left[I_d(t) - I_d^t\right] \\
			n_p(t+1) &= n_p(t) + \mu_n \left[V_p^t - \left( I_p(t) - \tilde{I}_p(t)\right)^2\right] \\
			n_d(t+1) &= n_d(t) + \mu_n \left[V_d^t - \left( I_d(t) - \tilde{I}_d(t)\right)^2\right] \\
			\tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_p(t+1) \\
			\tilde{I}_d(t+1) &= (1-\mu_{av})\tilde{I}_d(t) + \mu_{av} I_d(t+1)
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $I_p^t$, $I_d^t$, $V_p^t$ and $V_p^t$ define targets for the 
		temporal means and variances of $I_p$ and $I_d$. The dynamic variables 
		$\tilde{I}_p$ and $\tilde{I}_d$ are simply low-pass filtered running 
		averages of $I_p$ and $I_d$.

		A list of all parameter values is given in Table~\ref{tab:parameters}.

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{table}
			\begin{tabular}{ l | l || l | l }
				$\theta_{p0}$ & $0$ & $V_d^t$ & $0.25$ \\
				$\theta_{p1}$ & $-1$ & $\mu_b$ & $10^{-3}$ \\ 
				$\theta_{d}$ & $0$ & $\mu_n$ & $10^{-4}$ \\  
				$\alpha$ & $0.3$ & $\mu_{av}$ & $5 \cdot 10^{-3}$ \\   
				$\mu_w$ & $5 \cdot 10^{-5}$ & $I_p^t$ & $0$ \\
				$V_p^t$ & $0.25$ & $I_d^t$ & $0$  
			\end{tabular}
		\caption{Model parameters}
		\label{tab:parameters}
		\end{table}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		\section{Results}

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\includegraphics[width=\columnwidth]{pd_corr_scaling}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\includegraphics[width=\columnwidth]{distraction_classification_compare}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		\bibliography{Dendritic_Learning}

\end{document}

### 2021-03-23 code/plotting/comp_model.py

@@ -1,0 +1,59 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+import seaborn as sns
+sns.set()
+from stdParams import *
+plt.style.use('mpl_style.mplstyle')
+import os
+
+def s(x):
+    return 1./(1.+np.exp(-4.*g*x))
+    
+def phi(x,y):
+    return alpha * s(x)*(1.-s(y)) + s(x-thp)*s(y)
+
+alpha = 0.3
+thp = -1.
+g=1.
+
+p = np.linspace(-2.,2.,100)
+d = np.linspace(-2.,2.,100)
+
+p_ax = np.array(p) - (p[1] - p[0])/2.
+p_ax = np.append(p_ax,2.*p_ax[-1] - p_ax[-2])
+
+d_ax = np.array(d) - (d[1] - d[0])/2.
+d_ax = np.append(d_ax,2.*d_ax[-1] - d_ax[-2])
+
+P,D = np.meshgrid(p,d)
+
+fig, ax = plt.subplots(1,1,figsize=(5,4))
+
+pm = ax.pcolormesh(p_ax,d_ax,phi(P,D),rasterized=True)
+
+divider = make_axes_locatable(ax)
+cax = divider.append_axes("right", size="5%", pad="13%")
+
+#fig.add_axes(cax)
+fig.colorbar(pm,cax=cax)
+
+ax.set_xlabel(r'$I_{\rm p}$')
+ax.set_ylabel(r'$I_{\rm d}$')
+
+ax.plot([0.,0.],[-2.,2.],'--',c='w')
+ax.plot([-1.,-1.],[-2.,2.],'--',c='w')
+ax.plot([-2.,2.],[0.,0.],'--',c='w')
+
+ax.text(1.,-1.2,r'$\alpha$',color='w',fontsize=14)
+ax.text(-.15,2.15,r'$\theta_{p0}$',fontsize=14)
+ax.text(-1.15,2.15,r'$\theta_{p1}$',fontsize=14)
+ax.text(2.15,-0.1,r'$\theta_{d}$',fontsize=14)
+
+fig.tight_layout(pad=0.1)
+
+fig.savefig("../../../notes/figures/plot_comp_mod_marks.pdf",transparent=True)
+fig.savefig("../../../notes/figures/plot_comp_mod_marks.png",dpi=600,transparent=True)

### 2021-03-23 code/plotting/pd_corr_dimension.py

@@ -1,0 +1,34 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+plt.style.use('mpl_style.mplstyle')
+import os
+
+data = np.load(os.path.join(DATA_DIR,"distraction_dimension.npz"))
+
+rho = data["align_corrcoef"]
+dim = data["distract_dimension"]
+
+rho_mean = rho.mean(axis=1)
+rho_err = rho.std(axis=1)/rho.shape[1]**.5
+rho_top = rho_mean + rho_err
+rho_bottom = rho_mean - rho_err
+
+fig, ax = plt.subplots(1,1,figsize=(FIG_WIDTH,0.6*FIG_WIDTH))
+
+ax.fill_between(dim,rho_bottom[:,0],rho_top[:,0],alpha=0.5,color=COLORS[0])
+ax.fill_between(dim,rho_bottom[:,1],rho_top[:,1],alpha=0.5,color=COLORS[1])
+
+ax.plot(dim,rho_mean[:,0],c=COLORS[0])
+ax.plot(dim,rho_mean[:,1],c=COLORS[1])
+
+ax.set_ylabel(r'$\rho[I_p,I_d]$')
+ax.set_xlabel(r'$N$')
+
+fig.savefig(os.path.join(PLOT_DIR,"pd_corr_dimension.pdf"))
+
+plt.show()

### 2021-03-23 code/simulation/dimension_correlation.py

@@ -1,0 +1,134 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+N_sweep_distraction = 9
+N_samples = 2
+distract_scaling = 2.
+distract_dimension = np.arange(1,N_sweep_distraction+1)
+
+align_corrcoef = np.ndarray((N_sweep_distraction,N_samples,2))
+
+N_p = 10
+
+T = int(5e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+modes = ["comp","point"]
+
+for mode in tqdm(modes):
+    for s in tqdm(range(N_sweep_distraction),leave=False):
+        for i in tqdm(range(N_samples),leave=False):
+            
+            #generate random orthonormal basis via qr-decomposition
+            Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                        
+            sequences = np.random.rand(T,N_p)
+            
+            w_p = np.ones((T,1,N_p))
+            
+            n_d = np.ones((T))
+            n_p = np.ones((T))
+            b_d = np.zeros((T))
+            b_p = np.zeros((T))
+            
+            x_d = (Q[:,0] @ sequences.T).T
+            
+            #Transform x_p into orthogonal basis,
+            #Scale up distract_dimension[s] dimensions
+            #and transform back.
+            x_p = (Q.T @ sequences.T).T
+            x_p[:,1:distract_dimension[s]+1] *= distract_scaling
+            x_p = (Q @ x_p.T).T
+            
+            I_p = np.ndarray((T))
+            I_d = np.ndarray((T))
+            
+            x_p_av = np.ndarray((T,N_p))
+            
+            I_p_av = np.ndarray((T))
+            I_d_av = np.ndarray((T))
+            
+            y = np.ndarray((T))
+            
+            y_squ_av = np.ndarray((T))
+            y_av = np.ndarray((T))
+            
+            #### Init values
+            
+            w_p[0] /= np.linalg.norm(w_p[0])
+            
+            I_p[0] = n_p[0] * (w_p[0] @ x_p[0]) - b_p[0]
+            I_d[0] = n_d[0] * x_d[0] - b_d[0]
+            
[…]

### 2021-03-23 code/simulation/scaling_correlation.py

@@ -14,6 +14,7 @@ align_corrcoef = np.ndarray((N_sweep_distraction,N_samples,2))
 N_p = 10
 
 T = int(5e5)
+T_test = int(1e4)
 t_ax = np.arange(T)
 
 modes = ["comp","point"]
@@ -21,18 +22,9 @@ modes = ["comp","point"]
 for mode in tqdm(modes):
     for s in tqdm(range(N_sweep_distraction),leave=False):
         for i in tqdm(range(N_samples),leave=False):
-            a = np.random.rand(1,N_p)
-
-            a /= np.linalg.norm(a)
-
-            #####
-            # This just generates a random vector and subtracts the projection of the a vector,
-            # which makes vector a and vector a_orth orthogonal. a_orth is then normalized to unit length
-            # -> Gram Schmidt
-            a_orth = np.random.rand(1,N_p)
-            a_orth = a_orth - (a_orth @ a.T)*a/(a @ a.T)
-            a_orth /= np.linalg.norm(a_orth)
-            
+            #generate random orthonormal basis via qr-decomposition
+            Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                        
             sequences = np.random.rand(T,N_p)
             
             w_p = np.ones((T,1,N_p))
@@ -42,10 +34,14 @@ for mode in tqdm(modes):
             b_d = np.zeros((T))
             b_p = np.zeros((T))
             
-            x_d = (a @ sequences.T).T
+            x_d = (Q[:,0] @ sequences.T).T
             
-            x_p = np.array(sequences)
-            x_p = x_p + ((x_p @ a_orth.T)*(distract_scaling[s] - 1.)) @ a_orth
+            #Transform x_p into orthogonal basis,
+            #Scale up distract_dimension[s] dimensions
+            #and transform back.
+            x_p = (Q.T @ sequences.T).T
+            x_p[:,1] *= distract_scaling[s]
+            x_p = (Q @ x_p.T).T
             
             I_p = np.ndarray((T))
             I_d = np.ndarray((T))
@@ -83,7 +79,8 @@ for mode in tqdm(modes):
             for t in tqdm(range(1,T),disable=False,leave=False):
                 
[…]

### 2021-03-23 code/simulation/scaling_dimension_classification.py

@@ -1,0 +1,145 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(1.,6.,N_sweep_distraction_scaling)
+distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,2))
+
+N_p = 10
+N_out = 2
+
+T = int(2e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+stdMainDir = .25
+distMainDir = 2.
+
+modes = ["comp","point"]
+
+for mode in tqdm(modes):
+    for s in tqdm(range(N_sweep_distraction_scaling),leave=False):
+        for n in tqdm(range(N_sweep_distraction_dimension),leave=False):
+            for i in tqdm(range(N_samples),leave=False):
+                
+                
+                patterns = np.random.normal(0.,1.,(T,N_p))
+                patterns[:,0] *= stdMainDir
+                patterns[:,1:distract_dimension[n]+1] *= distract_scaling[s]
+                patterns[:,distract_dimension[n]+1:] *= 0.
+                patterns[:,0] += (1.*(np.random.rand(T) <= 0.5) - 0.5) * distMainDir
+
+                
+                w_p = np.ones((T,N_out,N_p))
+                
+                n_d = np.ones((T,N_out))
+                n_p = np.ones((T,N_out))
+                b_d = np.zeros((T,N_out))
+                b_p = np.zeros((T,N_out))
+                
+                x_d = np.ndarray((T,2))
+                x_d[:,0] = 1.*(patterns[:,0] < 0.)
+                x_d[:,1] = 1.*(patterns[:,0] > 0.)
+                
+                #generate random orthonormal basis via qr-decomposition
+                Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                #Transform x_p with orthogonal basis.
+                x_p = (Q @ patterns.T).T
+                
+                I_p = np.ndarray((T,N_out))
+                I_d = np.ndarray((T,N_out))
+                
[…]

### 2021-03-23 code/simulation/scaling_dimension_correlation.py

@@ -1,0 +1,138 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 2
+distract_scaling = np.linspace(1.,6.,N_sweep_distraction_scaling)
+distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
+
+align_corrcoef = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,2))
+
+N_p = 10
+
+T = int(5e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+modes = ["comp","point"]
+
+for mode in tqdm(modes):
+    for s in tqdm(range(N_sweep_distraction_scaling),leave=False):
+        for n in tqdm(range(N_sweep_distraction_dimension),leave=False):
+            for i in tqdm(range(N_samples),leave=False):
+                #generate random orthonormal basis via qr-decomposition
+                Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                            
+                sequences = np.random.rand(T,N_p)
+                
+                w_p = np.ones((T,1,N_p))
+                
+                n_d = np.ones((T))
+                n_p = np.ones((T))
+                b_d = np.zeros((T))
+                b_p = np.zeros((T))
+                
+                x_d = (Q[:,0] @ sequences.T).T
+                
+                #Transform x_p into orthogonal basis,
+                #Scale up distract_dimension[s] dimensions
+                #and transform back.
+                x_p = (Q.T @ sequences.T).T
+                x_p[:,1:distract_dimension[n]+1] *= distract_scaling[s]
+                x_p = (Q @ x_p.T).T
+                
+                I_p = np.ndarray((T))
+                I_d = np.ndarray((T))
+                
+                x_p_av = np.ndarray((T,N_p))
+                
+                I_p_av = np.ndarray((T))
+                I_d_av = np.ndarray((T))
+                
+                y = np.ndarray((T))
+                
+                y_squ_av = np.ndarray((T))
+                y_av = np.ndarray((T))
[…]

### 2021-03-23 code/stdParams.py

@@ -10,7 +10,7 @@ COLORS = plt.rcParams['axes.prop_cycle'].by_key()['color']
 DATA_DIR = os.path.join(os.path.dirname(__file__),"../data")
 PLOT_DIR = os.path.join(os.path.dirname(__file__),"../plots")
 
-FIG_WIDTH = 3.47
+FIG_WIDTH = 5.0
 
 thetap0 = 0.
 thetap1 = -1.

### 2021-03-29 notes/draft_article.tex
.

@@
 		such as a mean square error, based on the difference between 
-		basal and apical input, i.e. $I_p - I_d$. Theoretical work has 
+		basal and apical input, i.e. $I_{\rm p} - I_{\rm d}$. Theoretical work has 
 		investigated possible learning mechanisms
@@
 		simplified version of a phenomenological model proposed by \citet{Shai_2015}.
-		Denoting the input currents by $I_p$ (proximal) and $I_d$ (distal), 
+		Denoting the input currents by $I_{\rm p}$ (proximal) and $I_{\rm d}$ (distal), 
 		the model is written as
@@
 		\begin{split}
-		y\left(t\right) &= \sigma\left( I_p(t) - \theta_{p0} \right)
-		\left[1-\sigma\left(I_d(t) - \theta_d\right)\right] \\
-		&+ \alpha \sigma\left(I_d(t) - \theta_d \right)\sigma\left( I_p(t) - \theta_{p1} \right)
+		y\left(t\right) &= \sigma\left( I_{\rm p}(t) - \theta_{p0} \right)
+		\left[1-\sigma\left(I_{\rm d}(t) - \theta_d\right)\right] \\
+		&+ \alpha \sigma\left(I_{\rm d}(t) - \theta_d \right)\sigma\left( I_{\rm p}(t) - \theta_{p1} \right)
 		\end{split} \label{eq:comp_model}\\
@@
 		respect to proximal and distal input. Overall, this equation describes two distinct
-		regions of neural activation in the $(I_p, I_d)$-space which differ in their
+		regions of neural activation in the $(I_{\rm p}, I_{\rm d})$-space which differ in their
 		maximal firing rates, which are set to $1$ and $\alpha$, where $0 < \alpha < 1$.
@@
 			\caption{{\bf Two-compartment rate model.} Firing rate as a
-			function of proximal and distal input $I_p$ and $I_d$, 
+			function of proximal and distal input $I_{\rm p}$ and $I_{\rm d}$, 
 			see \eqref{eq:comp_model}. 
@@
 		\begin{equation}
-			y(t) = \sigma\left(I_p(t) + I_d(t) - \theta \right) \; .
+			y(t) = \sigma\left(I_{\rm p}(t) + I_{\rm d}(t) - \theta \right) \; .
 		\end{equation}
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		In all our numerical experiments, the apical input $I_d$ was generated
+		In all our numerical experiments, the apical input $I_{\rm d}$ was generated
 		``as is", meaning, it was not dynamically calculated as a superposition
@@
 		\begin{equation}
-			I_d(t) = n_d(t) x_d(t) - b_d(t) \; ,
+			I_{\rm d}(t) = n_d(t) x_d(t) - b_d(t) \; ,
 			\label{eq:I_d}
@@
 
-		Similarly, $I_p(t)$ was given by
+		Similarly, $I_{\rm p}(t)$ was given by
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 		\begin{equation}
-			I_p(t) = n_p(t) \sum_{i=1}^{N} x_{p,i}(t) w_i(t) - b_p(t) \; ,
+			I_{\rm p}(t) = n_p(t) \sum_{i=1}^{N} x_{p,i}(t) w_i(t) - b_p(t) \; ,
 		\end{equation}
@@
 		corresponding sequences and $w_i(t)$ the synaptic efficacies.
-		As for $I_d(t)$, $n_p(t)$ and $b_p(t)$ was used as a time dependent
+		As for $I_{\rm d}(t)$, $n_p(t)$ and $b_p(t)$ was used as a time dependent
 		scaling and bias.
@@
 		\begin{align}
-			b_p(t+1) &= b_p(t) + \mu_b \left[I_p(t) - I_p^t\right] \\
-			b_d(t+1) &= b_d(t) + \mu_b \left[I_d(t) - I_d^t\right] \\
-			n_p(t+1) &= n_p(t) + \mu_n \left[V_p^t - \left( I_p(t) - \tilde{I}_p(t)\right)^2\right] \\
-			n_d(t+1) &= n_d(t) + \mu_n \left[V_d^t - \left( I_d(t) - \tilde{I}_d(t)\right)^2\right] \\
-			\tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_p(t+1) \\
-			\tilde{I}_d(t+1) &= (1-\mu_{av})\tilde{I}_d(t) + \mu_{av} I_d(t+1)
+			b_p(t+1) &= b_p(t) + \mu_b \left[I_{\rm p}(t) - I_{\rm p}^t\right] \\
+			b_d(t+1) &= b_d(t) + \mu_b \left[I_{\rm d}(t) - I_{\rm d}^t\right] \\
+			n_p(t+1) &= n_p(t) + \mu_n \left[V_p^t - \left( I_{\rm p}(t) - \tilde{I}_p(t)\right)^2\right] \\
+			n_d(t+1) &= n_d(t) + \mu_n \left[V_d^t - \left( I_{\rm d}(t) - \tilde{I}_d(t)\right)^2\right] \\
+			\tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_{\rm p}(t+1) \\
+			\tilde{I}_d(t+1) &= (1-\mu_{av})\tilde{I}_d(t) + \mu_{av} I_{\rm d}(t+1)
 		\end{align}
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		Here, $I_p^t$, $I_d^t$, $V_p^t$ and $V_p^t$ define targets for the 
-		temporal means and variances of $I_p$ and $I_d$. The dynamic variables 
+		Here, $I_{\rm p}^t$, $I_{\rm d}^t$, $V_p^t$ and $V_p^t$ define targets for the 
+		temporal means and variances of $I_{\rm p}$ and $I_{\rm d}$. The dynamic variables 
 		$\tilde{I}_p$ and $\tilde{I}_d$ are simply low-pass filtered running 
-		averages of $I_p$ and $I_d$.
+		averages of $I_{\rm p}$ and $I_{\rm d}$.
 
@@
 				$\alpha$ & $0.3$ & $\mu_{av}$ & $5 \cdot 10^{-3}$ \\   
-				$\mu_w$ & $5 \cdot 10^{-5}$ & $I_p^t$ & $0$ \\
-				$V_p^t$ & $0.25$ & $I_d^t$ & $0$  
+				$\mu_w$ & $5 \cdot 10^{-5}$ & $I_{\rm p}^t$ & $0$ \\
+				$V_p^t$ & $0.25$ & $I_{\rm d}^t$ & $0$  
 			\end{tabular}
@@
 		\section{Results}
+		\label{sect:results}
+
+		\subsection{Increased Alignment between Basal and Apical Inputs}
+		\label{sect:alignment}
+
+		As a first test, we wanted to quantify the neuron's ability to 
+		align its basal input to the apical teaching signal.
+		To do so, we defined the pearson correlation coefficient
+		$\rho[I_{\rm p},I_{\rm d}]$ between the basal and apical input current
+		as our measure of interest. We determined this temporal
+		correlation coefficient after simulating all plasticity mechanisms
+		under stationary random input sequences with certain statistical
+		properties that shall be explained in the following.
+		As a starting point, we chose all $x_{p,i}(t)$ to be randomly
+		drawn from a uniform distribution, where $x_{p,i}(t) \in (0,1)$.
+		For $I_{\rm d}(t)$ to be fully `reconstructable' by the basal input, 
+		$x_d(t)$ had to be some linear combination of all $x_{p,i}(t)$.
+		Therefore, we chose $x_d(t) = \sum_{i=1}^N a_i x_{p,i}(t)$, where
+		$\mathbf{a}$ is a random vector with unit length.
+		Since we used a Hebbian learning scheme, we expected that 
+		the direction and magnitude of the principal components of 
+		the basal input would also significantly affect the outcome of
+		the experiment: A large variance in the basal input 
+		orthogonal to the `reconstruction vector' $\mathbf{a}$ 
+		should act as a distraction for the plasticity and reduce the
+		resulting alignment between $I_{\rm p}$ and $I_{\rm d}$. Therefore, we 
+		applied a transformation to the input sequences $x_{p,i}(t)$
+		that were parameterized by two quantities, a scaling factor $s$
+		and the dimension $N_{dist}$ of a subspace of the basal input space.
+		A set of $N_{dist}$ orthonormal basis vectors was randomly generated,
+		which were also orthogonal to $\mathbf{a}$. Withing this $N_{dist}$-dimensional
+		subspace, the input sequences $x_{p,i}(t)$ were then rescaled by the
+		factor $s$.
+
+		After both $x_{p,i}(t)$ and $x_d(t)$ were generated,
+		a simulation was run using all previously described plasticity
+		mechanisms until the dynamic variables reached a stationary state.
+		After this learning phase, another set of input sequences was generated
+		using the previously described protocol and $\rho[I_{\rm p},I_{\rm d}]$ was calculated.
+		Note that plasticity was turned off in this phase. This entire procedure
+		allowed us to calculate $\rho[I_{\rm p},I_{\rm d}]$ as a function of the distraction
+		parameters $s$ and $N_{dist}$. The results for both neuron models 
+		is shown in Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
+		basal inputs was $N=10$. One can observe a decorrelation transition for 
+		both models. However, the compartment model supports a significantly stronger
+		distraction in terms of the scaling factor $s$ as compared to the point model.
+		This was a first confirmation of our hypothesis that nonlinear interactions
+		between basal and apical input could improve learning guided by top-down signals.
 
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{pd_corr_scaling}
+			\includegraphics[width=\columnwidth]{corr_dimension_scaling}
+			\caption{{\bf Alignment between Basal and Apical Input.} Color
+			encodes the Pearson correlation $\rho[I_p,I_d]$ for different
+			number of orthogonal distraction directions $N_{dist}$ 
+			and the corresponding scaling faction $s$.}
+			\label{fig:corr_dimension_scaling}
 		\end{figure}
@@
 
+		\subsection{Supervised Learning in a Linear Classification Task}
+
+		Next, we investigated if the observed differences would also improve
+		the performance in an actual supervised learning task.
+		For this purpose, we constructed presynaptic basal input as illustrated
+		in Fig.~\ref{fig:illustration_classification}. Written in vector form,
+		each sample from the basal input was generated from the following expression:
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{equation}
+			\mathbf{x}_p(t) = \mathbf{b} + \mathbf{a}\left(c(t) + \sigma_a \zeta_a(t) \right) 
+			+ s \cdot \sum_{i=1}^{N_{dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
+		\end{equation}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		Here, $\mathbf{b}$ is a random vector drawn uniformly from
+		$(0,1)^N$, $\mathbf{a}$ is random unit vector as introduced in 
+		Section~\ref{sect:alignment}, $c(t)$ is a binary variable drawn 
+		from $\{-0.5,0.5\}$ with equal probability and $\zeta_a(t)$ and the
+		$\zeta_{dist,i}(t)$ are independent random Gaussian variables.
+		Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
+		orthogonal basis of $N_{dist}$ unit vectors which---just as in 
+		Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
+		As indicated by the time dependence, the Gaussian and binary random
+		variables were drawn for each time step. The vectors
+		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ were generated
+		once before the beginning of a simulation run.
+
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{distraction_classification_compare}
+			\centering
+			\includegraphics[width=0.75\columnwidth]{illustration_classification}
+			\caption{{\bf Input Space for a Simple Linear Classification Task.}
+			Two clusters of presynaptic basal activity were generated from 
+			multivariate Gaussian distributions. Here, $s$ denotes the standard
+			deviation orthogonal to the normal vector of the classification
+			hyperplane.}
+			\label{fig:illustration_classification}
 		\end{figure}

### 2021-03-29 code/plotting/classification_dimension_scaling.py

@@ -1,0 +1,49 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+
+data = np.load(os.path.join(DATA_DIR,"classification_scaling_dimension.npz"))
+
+perf = data["perf"]
+dim = data["distract_dimension"]
+s = data["distract_scaling"]
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = perf.shape[2]
+
+perf_mean = perf.mean(axis=2)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
+pc0 = ax[0].pcolormesh(s_ax,dim_ax,perf_mean[:,:,0].T,rasterized=True,vmin=0.5,vmax=1.)
+pc1 = ax[1].pcolormesh(s_ax,dim_ax,perf_mean[:,:,1].T,rasterized=True,vmin=0.5,vmax=1.)
+
+ax[0].set_xlabel(r'$s$')
+ax[1].set_xlabel(r'$s$')
+
+ax[0].set_ylabel(r'$N_{dist}$')
+ax[1].set_ylabel(r'$N_{dist}$')
+
+ax[0].set_xlim(right=5.)
+ax[1].set_xlim(right=5.)
+
+ax[0].set_title("Compartment Model",loc="right")
+ax[1].set_title("Point Model",loc="right")
+
+plt.colorbar(pc1,cax=ax[2])
+
+fig.tight_layout(pad=0.2)
+
+fig.savefig(os.path.join(PLOT_DIR,"classification_dimension_scaling.pdf"))
+fig.savefig(os.path.join(PLOT_DIR,"classification_dimension_scaling.png"),dpi=600)
+
+plt.show()

### 2021-03-29 code/plotting/corr_dimension_scaling.py

@@ -1,0 +1,49 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+
+data = np.load(os.path.join(DATA_DIR,"distraction_scaling_dimension.npz"))
+
+rho = data["align_corrcoef"]
+dim = data["distract_dimension"]
+s = data["distract_scaling"]
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = rho.shape[2]
+
+rho_mean = rho.mean(axis=2)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
+pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
+pc1 = ax[1].pcolormesh(s_ax,dim_ax,rho_mean[:,:,1].T,rasterized=True,vmin=0.,vmax=1.)
+
+ax[0].set_xlabel(r'$s$')
+ax[1].set_xlabel(r'$s$')
+
+ax[0].set_ylabel(r'$N_{dist}$')
+ax[1].set_ylabel(r'$N_{dist}$')
+
+ax[0].set_xlim(right=4.)
+ax[1].set_xlim(right=4.)
+
+ax[0].set_title("Compartment Model",loc="right")
+ax[1].set_title("Point Model",loc="right")
+
+plt.colorbar(pc1,cax=ax[2])
+
+fig.tight_layout(pad=0.2)
+
+fig.savefig(os.path.join(PLOT_DIR,"corr_dimension_scaling.pdf"))
+fig.savefig(os.path.join(PLOT_DIR,"corr_dimension_scaling.png"),dpi=600)
+
+plt.show()

### 2021-03-29 code/simulation/scaling_dimension_correlation.py

@@ -23,6 +23,9 @@ t_ax = np.arange(T)
 
 modes = ["comp","point"]
 
+adapt_rule = "hebb"
+#adapt_rule = "fisher"
+
 for mode in tqdm(modes):
     for s in tqdm(range(N_sweep_distraction_scaling),leave=False):
         for n in tqdm(range(N_sweep_distraction_dimension),leave=False):
@@ -83,13 +86,12 @@ for mode in tqdm(modes):
 
                 for t in tqdm(range(1,T),disable=False,leave=False):
                     
-                    if(mode=="comp"):
-                        #w_p[t] = w_p[t-1] + mu_w * (x_p[t-1] - x_p_av[t-1]) * y[t-1]*(y[t-1]-thetay)
+                    if(adapt_rule=="hebb"):
                         w_p[t] = w_p[t-1] + mu_w * (x_p[t-1] - x_p_av[t-1]) * (y[t-1]-y_av[t-1])
-                    else:
-                        #w_p[t] = w_p[t-1] + mu_w * (x_p[t-1] - x_p_av[t-1]) * y[t-1]*(y[t-1]-y_squ_av[t-1])
-                        w_p[t] = w_p[t-1] + mu_w * (x_p[t-1] - x_p_av[t-1]) * (y[t-1]-y_av[t-1])
-                    
+                    elif(adpat_rule=="fisher"):
+                        pass
+                        #w_p[t] = w_p[t-1] + mu_w * 
+                        
                     w_p[t] /= np.linalg.norm(w_p[t])
                     
                     b_p[t] = b_p[t-1] + mu_b * (I_p[t-1] - I_pt)

### 2021-03-29 code/src/plottools.py

@@ -1,0 +1,7 @@
+#! /usr/bin/env python3
+import numpy as np
+
+def gen_mesh_ax(x):
+    x_mesh = np.append(x,2.*x[-1] - x[-2])
+    x_mesh -= .5 * (x_mesh[1] - x_mesh[0])
+    return x_mesh

### 2021-03-29 code/stdParams.py

@@ -10,7 +10,7 @@ COLORS = plt.rcParams['axes.prop_cycle'].by_key()['color']
 DATA_DIR = os.path.join(os.path.dirname(__file__),"../data")
 PLOT_DIR = os.path.join(os.path.dirname(__file__),"../plots")
 
-FIG_WIDTH = 5.0
+FIG_WIDTH = 5.5
 
 thetap0 = 0.
 thetap1 = -1.

### 2021-03-30 notes/draft_article.tex
.#

@@
 		\subsection{Supervised Learning in a Linear Classification Task}
+		\label{sect:classification}
 
@@
 		For this purpose, we constructed presynaptic basal input as illustrated
-		in Fig.~\ref{fig:illustration_classification}. Written in vector form,
+		in Fig.~\ref{fig:illustration_classification}.
+		Written in vector form,
 		each sample from the basal input was generated from the following expression:
@@
 		from $\{-0.5,0.5\}$ with equal probability and $\zeta_a(t)$ and the
-		$\zeta_{dist,i}(t)$ are independent random Gaussian variables.
+		$\zeta_{dist,i}(t)$ are independent random Gaussian variables with 
+		zero mean and unit variance. 
+		Hence, $\sigma_a$ simply denotes the standard deviation of each Gaussian
+		cluster along the direction of the normal vector $\mathbf{a}$ and was
+		set to $\sigma_a = 0.25$. 
 		Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
@@
 		Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
+		The free parameter $s$ parameterized
+		the standard deviation along this subspace orthogonal to $\mathbf{a}$.
 		As indicated by the time dependence, the Gaussian and binary random
@@
 		once before the beginning of a simulation run.
-
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+		For the classification task, we created two output neurons receiving
+		the same basal presynaptic input, while the top-down input encoded
+		the correct linear classification in a one-hot scheme, that is
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{align}
+			x_{d,0}(t) &= 1 - \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \\
+			x_{d,1}(t) &= \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \; ,
+		\end{align}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		where $\Theta(x)$ is the Heaviside step function.
+
+		As in the previous experiment, we ran a full simulation until 
+		all dynamic variables reached a stationary state. After this,
+		a test run without plasticity and with the apical input turned off 
+		was used to evaluate the classification	performance. For each sample,
+		the index of the neuron with the highest activity was used
+		as the predicted class. Accuracy was then calculated as the fraction
+		of correctly classified samples.
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{figure}
+			\includegraphics[width=\columnwidth]{classification_dimension_scaling}
+			\caption{{\bf Binary Classification Accuracy.}
+			Fraction of correctly classified patterns as illustrated in
+			Fig.~\ref{fig:illustration_classification}, see 
+			Section~\ref{sect:classification}.¸}
+		\label{fig:classification_accuracy}
+		\end{figure}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+		The resulting accuracy as a function of $N_{dist}$ and $s$
+		is shown in Fig.~\ref{fig:classification_accuracy}.
+		Albeit differences were small, the compartment model did
+		show a better overall accuracy for the tested parameter range.
+
+		\subsection{Non-Hebbian Learning Rules}
+		\label{sect:non-hebbian}
+
+		To be continued, see Fig.~\ref{fig:classification_accuracy_bcm}.
+
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{figure}
+			\includegraphics[width=\columnwidth]{classification_dimension_scaling_bcm}
+			\caption{{\bf Binary Classification Accuracy, BCM Rule.}
+				Fraction of correctly classified patterns as illustrated in
+				Fig.~\ref{fig:illustration_classification}, after training with
+				a BCM-like learning rule.¸}
+			\label{fig:classification_accuracy_bcm}
+		\end{figure}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+		\section{Discussion}
+
 		\bibliography{Dendritic_Learning}

### 2021-03-30 code/plotting/classification_dimension_scaling_bcm.py

@@ -1,0 +1,49 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+
+data = np.load(os.path.join(DATA_DIR,"classification_scaling_dimension_bcm.npz"))
+
+perf = data["perf"]
+dim = data["distract_dimension"]
+s = data["distract_scaling"]
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = perf.shape[2]
+
+perf_mean = perf.mean(axis=2)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
+pc0 = ax[0].pcolormesh(s_ax,dim_ax,perf_mean[:,:,0].T,rasterized=True,vmin=0.5,vmax=1.)
+pc1 = ax[1].pcolormesh(s_ax,dim_ax,perf_mean[:,:,1].T,rasterized=True,vmin=0.5,vmax=1.)
+
+ax[0].set_xlabel(r'$s$')
+ax[1].set_xlabel(r'$s$')
+
+ax[0].set_ylabel(r'$N_{dist}$')
+ax[1].set_ylabel(r'$N_{dist}$')
+
+#ax[0].set_xlim(right=5.)
+#ax[1].set_xlim(right=5.)
+
+ax[0].set_title("Compartment Model",loc="right")
+ax[1].set_title("Point Model",loc="right")
+
+plt.colorbar(pc0,cax=ax[2])
+
+fig.tight_layout(pad=0.2)
+
+fig.savefig(os.path.join(PLOT_DIR,"classification_dimension_scaling_bcm.pdf"))
+fig.savefig(os.path.join(PLOT_DIR,"classification_dimension_scaling_bcm.png"),dpi=600)
+
+plt.show()

### 2021-03-30 code/simulation/scaling_dimension_classification_bcm.py

@@ -11,6 +11,7 @@ N_samples = 1
 distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
 distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
 
+from multiprocessing import Pool
 
 
 N_p = 10
@@ -31,6 +32,144 @@ perf = np.ndarray((N_sweep_distraction_scaling,
                     N_sweep_distraction_dimension,
                     N_samples,len(modes)))
 
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+
+def run_sim(arglist):
+    
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    patterns = np.random.normal(0.,1.,(T,N_p))
+    patterns[:,0] *= stdMainDir
+    patterns[:,1:N_distr+1] *= s
+    patterns[:,N_distr+1:] *= 0.
+    patterns[:,0] += (1.*(np.random.rand(T) <= 0.5) - 0.5) * distMainDir
+
+    
+    w_p = np.ones((T,N_out,N_p))
+    
+    n_d = np.ones((T,N_out))
+    n_p = np.ones((T,N_out))
+    b_d = np.zeros((T,N_out))
+    b_p = np.zeros((T,N_out))
+    
+    x_d = np.ndarray((T,2))
+    x_d[:,0] = 1.*(patterns[:,0] < 0.)
+    x_d[:,1] = 1.*(patterns[:,0] > 0.)
+    
+    #generate random orthonormal basis via qr-decomposition
+    Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+    #Transform x_p with orthogonal basis.
+    x_p = (Q @ patterns.T).T
+    
+    I_p = np.ndarray((T,N_out))
+    I_d = np.ndarray((T,N_out))
+    
+    x_p_av = np.ndarray((T,N_p))
+    
+    I_p_av = np.ndarray((T,N_out))
+    I_d_av = np.ndarray((T,N_out))
+    
+    y = np.ndarray((T,N_out))
+    
+    y_squ_av = np.ndarray((T,N_out))
+    y_av = np.ndarray((T,N_out))
+    
[…]

### 2021-04-06 notes/draft_article.tex
.

@@
 
-		In our work, we combined a phenomenological model predicting the output
+		In our work, we combined a phenomenological model that generates the output
 		firing rate as a function of two streams of synaptic input (subsuming basal and apical input)
-		with a BCM-like plasticity rule on basal synapses. LTP was induced in our model 
-		if the postsynaptic activity surpassed a threshold level that was only achievable if 
-		both basal and apical inputs were present. Therefore, we hypothesized that 
+		with Hebbian as well as BCM-like plasticity rules on basal synapses. We hypothesized that 
 		this combination of neural activation and plasticity rules would lead to an
@@
 		learning rule. Thus, we also tested our model in a simple linear 
-		supervised classification task and compared it to the performance of simple
-		point neuron combined with a BCM-plasticity rule on the basal inputs.
+		supervised classification task and compared it to the performance of a simple
+		point neuron equipped with similar plasticity rules.
 
@@
 
-		We implemented a BCM-like plasticity rule for the basal 
+		We implemented a Hebbian plasticity rule for the basal 
 		synaptic weights given by the following update equation:
@@
 			w_i(t+1) &= w_i(t) + \mu_w \left(x_{p,i}(t) - \tilde{x}_{p,i}(t)\right)
-			y(t)\left[y(t) - \theta_y \right]
+			\left(y(t) - \tilde{y} \right)
 			\label{eq:bcm_plast} \\
-			\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1)
+			\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1) \\
+			\tilde{y}(t+1) &= (1-\mu_{av})\tilde{y}(t) + \mu_{av}y(t+1)
 		\end{align}
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		While the classical BCM-rule dynamically modifies the 
-		threshold $\theta_y$ to stabilize weights, we used a
+		Additionally, we used a
 		synaptic normalization constraint
@@
 			w_i(t) \rightarrow \frac{w_i(t)}{\lVert \mathbf{w}(t)\rVert}
+			\label{eq:weight_norm}
 		\end{equation}
@@
 		in each time step, where $\lVert \mathbf{w}(t)\rVert$ denotes the Euclidean
-		norm of the synaptic weight vector. The threshold $\theta_y$ was
-		chosen such that LTP is present if neural activity is in the
-		high-activity regime (yellow area in Fig.~\ref{fig:comp_model}), while
-		LTD is present in the low-activity regime 
-		(given by $\alpha$, see Fig.~\ref{fig:comp_model}). Hence, we set
-		$\theta_y = (1+\alpha)/2$.
+		norm of the synaptic weight vector. 
+%		The threshold $\theta_y$ was
+%		chosen such that LTP is present if neural activity is in the
+%		high-activity regime (yellow area in Fig.~\ref{fig:comp_model}), while
+%		LTD is present in the low-activity regime 
+%		(given by $\alpha$, see Fig.~\ref{fig:comp_model}). Hence, we set
+%		$\theta_y = (1+\alpha)/2$.
 
 		For comparative reasons, the point neuron model was equipped with the
-		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}, 
-		except that the threshold $\theta_y$ was dynamically adjusted as
-		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		\begin{equation}
-			\theta_y(t+1) = \left(1-\mu_{av}\right)\theta_y(t) + \mu_{av} y^2(t) \; , 
-		\end{equation}
-		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		which is an implementation of a running average of the square activity,
-		as used in the classic BCM-rule.
+		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}. 
+%		except that the threshold $\theta_y$ was dynamically adjusted as
+%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%		\begin{equation}
+%			\theta_y(t+1) = \left(1-\mu_{av}\right)\theta_y(t) + \mu_{av} y^2(t) \; , 
+%		\end{equation}
+%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%		which is an implementation of a running average of the square activity,
+%		as used in the classic BCM-rule.
 
@@
 
-		To be continued, see Fig.~\ref{fig:classification_accuracy_bcm}.
+		Instead of Hebbian learning, we also considered a BCM-like
+		learning rule for the basal weights \citep{Bienenstock1982,Intrator1992}.
+		The form of the BCM-rule we consider here reads
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{equation}
+		\Delta w_i \propto y\left(y - \theta_M\right) x_i - \epsilon w_i \; ,
+		\end{equation}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		where $\theta_M$ is a threshold defining a transition from LTP to LTD and
+		$\epsilon$ is an optional decay term on the weights.
+		In the variant introduced by \citet{Law1994}, the sliding threshold is simply
+		the temporal average of the squared neural activity, 
+		$\theta_M = \langle y^2 \rangle$. In practice, this would be calculated
+		as a running average, thereby preventing the weights from growing 
+		indefinitely.
+
+		However, for our compartment model, we chose to explicitly set the
+		threshold to be the mean value between the high- and low-activity regime
+		in our compartment model, i.e.\ $\theta_M = (1+\alpha)/2$. By doing so, LTP was
+		preferably induced if both basal and apical input was present at the same
+		time. Furthermore, instead of the weight decay term, we chose to keep
+		the weight normalization as introduced in \eqref{eq:weight_norm}.
+		Obviously, for the point model, the reasoning behind our choice of
+		$\theta_M$ did not apply. Still, to provide some level of comparability,
+		we also ran simulations with a point model where the sliding threshold was
+		calculated as a running average of $y^2$. Furthermore, we did not use
+		weight normalization in this case, but chose to use a small weight decay term
+		with $\epsilon = 0.1$. The results are shown in 
+		Fig.~\ref{fig:classification_accuracy_bcm} (classification task) and 
+		Fig.~\ref{fig:classification_accuracy_bcm} (Basal-Apical alignment). 
+		While the accuracy of the classification for the point model
+		was at most comparable to Hebbian learning, the BCM-like rule for the 
+		compartment model significantly increased the accuracy for the tested
+		parameter range (compare Fig.~\ref{fig:classification_accuracy}).
+
+		Likewise, the compartment model also significantly benefited from
+		the BCM rule in terms of basal-apical alignment, while only marginal
+		improvements could be observed for the point model
+		(compare Fig.~\ref{fig:corr_dimension_scaling_bcm} 
+		with Fig.~\ref{fig:corr_dimension_scaling}).
 
@@
 
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{figure}
+			\includegraphics[width=\columnwidth]{corr_dimension_scaling_bcm}
+			\caption{{\bf Alignment between Basal and Apical Input, BCM Rule.}
+				Color encodes the Pearson correlation $\rho[I_p,I_d]$ for different
+				number of orthogonal distraction directions $N_{dist}$ 
+				and the corresponding scaling faction $s$ after training
+				with a BCM-like rule. Compare to Fig.~\ref{fig:corr_dimension_scaling}.}
+			\label{fig:corr_dimension_scaling_bcm}
+		\end{figure}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 		\section{Discussion}
+
+		To be continued
 

### 2021-04-06 code/plotting/classification_correlation_dimension_scaling.py

@@ -1,0 +1,103 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling/"+file)))
+
+perf = []
+Ip = []
+labels = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    Ip.append(dat["I_p_samples"])
+    labels.append(dat["labels_samples"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(Ip[k].shape,Ip[k-1].shape))):
+        print("Ip arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(labels[k].shape,labels[k-1].shape))):
+        print("labels arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+Ip = np.array(Ip)
+labels = np.array(labels)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = Ip.shape[0]*Ip.shape[3]
+
+
+Ip_flatten = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],Ip.shape[5],n_total_samples))
+labels_flatten = np.ndarray((labels.shape[1],labels.shape[2],labels.shape[4],labels.shape[5],n_total_samples))
+rho = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],n_total_samples))
+
+for k in range(Ip.shape[0]):
[…]

### 2021-04-06 code/plotting/classification_correlation_dimension_scaling_bcm.py

@@ -1,0 +1,103 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm/"+file)))
+
+perf = []
+Ip = []
+labels = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    Ip.append(dat["I_p_samples"])
+    labels.append(dat["labels_samples"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(Ip[k].shape,Ip[k-1].shape))):
+        print("Ip arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(labels[k].shape,labels[k-1].shape))):
+        print("labels arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+Ip = np.array(Ip)
+labels = np.array(labels)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = Ip.shape[0]*Ip.shape[3]
+
+
+Ip_flatten = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],Ip.shape[5],n_total_samples))
+labels_flatten = np.ndarray((labels.shape[1],labels.shape[2],labels.shape[4],labels.shape[5],n_total_samples))
+rho = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],n_total_samples))
+
+for k in range(Ip.shape[0]):
[…]

### 2021-04-06 code/plotting/classification_dimension_scaling.py

@@ -8,19 +8,55 @@ from stdParams import *
 from src.plottools import gen_mesh_ax
 plt.style.use('mpl_style.mplstyle')
 import os
+import sys
 
-data = np.load(os.path.join(DATA_DIR,"classification_scaling_dimension.npz"))
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling"))
 
-perf = data["perf"]
-dim = data["distract_dimension"]
-s = data["distract_scaling"]
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling/"+file)))
+
+perf = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(perf[k].shape,perf[k-1].shape))):
+        print("perf arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = perf.shape[0]*perf.shape[3]
+
+perf_flatten = np.ndarray((perf.shape[1],perf.shape[2],perf.shape[4],n_total_samples))
+
+for k in range(perf.shape[0]):
+    for l in range(perf.shape[3]):
+        perf_flatten[:,:,:,k*perf.shape[3]+l] = perf[k,:,:,l,:]
+
+perf = perf_flatten
 
 dim_ax = gen_mesh_ax(dim)
 s_ax = gen_mesh_ax(s)
 
 n_sweep = perf.shape[2]
 
-perf_mean = perf.mean(axis=2)
+perf_mean = perf.mean(axis=3)
 
 fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
 
@@ -33,8 +69,8 @@ ax[1].set_xlabel(r'$s$')
 ax[0].set_ylabel(r'$N_{dist}$')
[…]

### 2021-04-06 code/plotting/classification_dimension_scaling_bcm.py

@@ -8,19 +8,55 @@ from stdParams import *
 from src.plottools import gen_mesh_ax
 plt.style.use('mpl_style.mplstyle')
 import os
+import sys
 
-data = np.load(os.path.join(DATA_DIR,"classification_scaling_dimension_bcm.npz"))
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm"))
 
-perf = data["perf"]
-dim = data["distract_dimension"]
-s = data["distract_scaling"]
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm/"+file)))
+
+perf = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(perf[k].shape,perf[k-1].shape))):
+        print("perf arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = perf.shape[0]*perf.shape[3]
+
+perf_flatten = np.ndarray((perf.shape[1],perf.shape[2],perf.shape[4],n_total_samples))
+
+for k in range(perf.shape[0]):
+    for l in range(perf.shape[3]):
+        perf_flatten[:,:,:,k*perf.shape[3]+l] = perf[k,:,:,l,:]
+
+perf = perf_flatten
 
 dim_ax = gen_mesh_ax(dim)
 s_ax = gen_mesh_ax(s)
 
 n_sweep = perf.shape[2]
 
-perf_mean = perf.mean(axis=2)
+perf_mean = perf.mean(axis=3)
 
 fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
 

### 2021-04-06 code/plotting/corr_dimension_scaling.py

@@ -1,49 +1,0 @@
-#! /usr/bin/env python3
-
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-sns.set()
-from stdParams import *
-from src.plottools import gen_mesh_ax
-plt.style.use('mpl_style.mplstyle')
-import os
-
-data = np.load(os.path.join(DATA_DIR,"distraction_scaling_dimension.npz"))
-
-rho = data["align_corrcoef"]
-dim = data["distract_dimension"]
-s = data["distract_scaling"]
-
-dim_ax = gen_mesh_ax(dim)
-s_ax = gen_mesh_ax(s)
-
-n_sweep = rho.shape[2]
-
-rho_mean = rho.mean(axis=2)
-
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
-
-pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
-pc1 = ax[1].pcolormesh(s_ax,dim_ax,rho_mean[:,:,1].T,rasterized=True,vmin=0.,vmax=1.)
-
-ax[0].set_xlabel(r'$s$')
-ax[1].set_xlabel(r'$s$')
-
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
-
-ax[0].set_xlim(right=4.)
-ax[1].set_xlim(right=4.)
-
-ax[0].set_title("Compartment Model",loc="right")
-ax[1].set_title("Point Model",loc="right")
-
-plt.colorbar(pc1,cax=ax[2])
-
-fig.tight_layout(pad=0.2)
-
-fig.savefig(os.path.join(PLOT_DIR,"corr_dimension_scaling.pdf"))
-fig.savefig(os.path.join(PLOT_DIR,"corr_dimension_scaling.png"),dpi=600)
-
-plt.show()

### 2021-04-06 code/plotting/correlation_dimension_scaling.py

@@ -1,0 +1,87 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"correlation_dimension_scaling"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"correlation_dimension_scaling/"+file)))
+
+rho = []
+dim = []
+s = []
+
+for dat in data:
+    rho.append(dat["align_corrcoef"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(rho)):
+    if(not(np.array_equal(rho[k].shape,rho[k-1].shape))):
+        print("corr. arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+rho = np.array(rho)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = rho.shape[0]*rho.shape[3]
+
+rho_flatten = np.ndarray((rho.shape[1],rho.shape[2],rho.shape[4],n_total_samples))
+
+for k in range(rho.shape[0]):
+    for l in range(rho.shape[3]):
+        rho_flatten[:,:,:,k*rho.shape[3]+l] = rho[k,:,:,l,:]
+
+rho = rho_flatten
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = rho.shape[2]
+
+rho_mean = rho.mean(axis=3)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
+pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
[…]

### 2021-04-06 code/plotting/correlation_dimension_scaling_bcm.py

@@ -1,0 +1,87 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"correlation_dimension_scaling_bcm"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"correlation_dimension_scaling_bcm/"+file)))
+
+rho = []
+dim = []
+s = []
+
+for dat in data:
+    rho.append(dat["align_corrcoef"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(rho)):
+    if(not(np.array_equal(rho[k].shape,rho[k-1].shape))):
+        print("corr. arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+rho = np.array(rho)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = rho.shape[0]*rho.shape[3]
+
+rho_flatten = np.ndarray((rho.shape[1],rho.shape[2],rho.shape[4],n_total_samples))
+
+for k in range(rho.shape[0]):
+    for l in range(rho.shape[3]):
+        rho_flatten[:,:,:,k*rho.shape[3]+l] = rho[k,:,:,l,:]
+
+rho = rho_flatten
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = rho.shape[2]
+
+rho_mean = rho.mean(axis=3)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
[…]

### 2021-04-06 code/simulation/classification_dimension_scaling.py

@@ -1,0 +1,187 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
+distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
+
+N_sweep_distraction_dimension = distract_dimension.shape[0]
+
+N_p = 10
+N_out = 2
+
+T = int(2e5)
+T_test = int(1e4)
+T_sample = int(1e3)
+t_ax = np.arange(T)
+
+stdMainDir = .25
+distMainDir = 2.
+
+modes = ["comp","point"]
+#modes = ["comp_bcm","point_bcm"]
+#modes = ["comp"]
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes)))
+                    
+I_p_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample,N_out))
+                    
+labels_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample))
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+def run_sim(arglist):
+    
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    patterns = np.random.normal(0.,1.,(T,N_p))
+    patterns[:,0] *= stdMainDir
+    patterns[:,1:N_distr+1] *= s
+    patterns[:,N_distr+1:] *= 0.
[…]

### 2021-04-06 code/simulation/classification_dimension_scaling_bcm.py

@@ -1,0 +1,196 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
+distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
+
+
+
+N_p = 10
+N_out = 2
+
+T = int(2e5)
+T_test = int(1e4)
+T_sample = int(1e3)
+t_ax = np.arange(T)
+
+stdMainDir = .25
+distMainDir = 2.
+
+modes = ["comp","point"]
+#modes = ["comp_bcm","point_bcm"]
+#modes = ["point"]
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes)))
+
+I_p_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample,N_out))
+                    
+labels_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample))
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+
+def run_sim(arglist):
+    
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    patterns = np.random.normal(0.,1.,(T,N_p))
+    patterns[:,0] *= stdMainDir
+    patterns[:,1:N_distr+1] *= s
+    patterns[:,N_distr+1:] *= 0.
+    patterns[:,0] += (1.*(np.random.rand(T) <= 0.5) - 0.5) * distMainDir
+
+    
[…]

### 2021-04-06 code/simulation/correlation_dimension_scaling.py

@@ -1,0 +1,175 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 2
+distract_scaling = np.linspace(1.,6.,N_sweep_distraction_scaling)
+distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
+
+align_corrcoef = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,2))
+
+N_p = 10
+
+T = int(5e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+modes = ["comp","point"]
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes)))
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+def run_sim(arglist):
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    #generate random orthonormal basis via qr-decomposition
+    Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                
+    sequences = np.random.rand(T,N_p)
+    
+    w_p = np.ones((T,1,N_p))
+    
+    n_d = np.ones((T))
+    n_p = np.ones((T))
+    b_d = np.zeros((T))
+    b_p = np.zeros((T))
+    
+    x_d = (Q[:,0] @ sequences.T).T
+    
+    #Transform x_p into orthogonal basis,
+    #Scale up distract_dimension[s] dimensions
+    #and transform back.
+    x_p = (Q.T @ sequences.T).T
+    x_p[:,1:N_distr+1] *= s
+    x_p = (Q @ x_p.T).T
+    
[…]

### 2021-04-06 code/simulation/correlation_dimension_scaling_bcm.py

@@ -1,0 +1,181 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 2
+distract_scaling = np.linspace(1.,6.,N_sweep_distraction_scaling)
+distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
+
+align_corrcoef = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,2))
+
+N_p = 10
+
+T = int(5e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+modes = ["comp","point"]
+#modes = ["point"]
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+def run_sim(arglist):
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    #generate random orthonormal basis via qr-decomposition
+    Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                
+    sequences = np.random.rand(T,N_p)
+    
+    w_p = np.ones((T,1,N_p))
+    
+    n_d = np.ones((T))
+    n_p = np.ones((T))
+    b_d = np.zeros((T))
+    b_p = np.zeros((T))
+    
+    x_d = (Q[:,0] @ sequences.T).T
+    
+    #Transform x_p into orthogonal basis,
+    #Scale up distract_dimension[s] dimensions
+    #and transform back.
+    x_p = (Q.T @ sequences.T).T
+    x_p[:,1:N_distr+1] *= s
+    x_p = (Q @ x_p.T).T
+    
+    I_p = np.ndarray((T))
+    I_d = np.ndarray((T))
+    
+    x_p_av = np.ndarray((T,N_p))
+    
[…]

### 2021-04-06 code/simulation/scaling_dimension_classification.py

@@ -1,148 +1,0 @@
-#! /usr/bin/env python3
-
-import numpy as np
-from stdParams import *
-from src.neuronmodel import *
-from tqdm import tqdm
-
-N_sweep_distraction_scaling = 20
-N_sweep_distraction_dimension = 9
-N_samples = 1
-distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
-distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
-
-
-
-N_p = 10
-N_out = 2
-
-T = int(2e5)
-T_test = int(1e4)
-t_ax = np.arange(T)
-
-stdMainDir = .25
-distMainDir = 2.
-
-modes = ["comp","point"]
-#modes = ["comp_bcm","point_bcm"]
-#modes = ["point"]
-
-perf = np.ndarray((N_sweep_distraction_scaling,
-                    N_sweep_distraction_dimension,
-                    N_samples,len(modes)))
-
-for mode in tqdm(modes):
-    for s in tqdm(range(N_sweep_distraction_scaling),leave=False):
-        for n in tqdm(range(N_sweep_distraction_dimension),leave=False):
-            for i in tqdm(range(N_samples),leave=False):
-                
-                
-                patterns = np.random.normal(0.,1.,(T,N_p))
-                patterns[:,0] *= stdMainDir
-                patterns[:,1:distract_dimension[n]+1] *= distract_scaling[s]
-                patterns[:,distract_dimension[n]+1:] *= 0.
-                patterns[:,0] += (1.*(np.random.rand(T) <= 0.5) - 0.5) * distMainDir
-
-                
-                w_p = np.ones((T,N_out,N_p))
-                
-                n_d = np.ones((T,N_out))
-                n_p = np.ones((T,N_out))
-                b_d = np.zeros((T,N_out))
-                b_p = np.zeros((T,N_out))
-                
-                x_d = np.ndarray((T,2))
-                x_d[:,0] = 1.*(patterns[:,0] < 0.)
-                x_d[:,1] = 1.*(patterns[:,0] > 0.)
-                
-                #generate random orthonormal basis via qr-decomposition
-                Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
-                #Transform x_p with orthogonal basis.
-                x_p = (Q @ patterns.T).T
-                
-                I_p = np.ndarray((T,N_out))
[…]

### 2021-04-06 code/simulation/scaling_dimension_classification_bcm.py

@@ -1,290 +1,0 @@
-#! /usr/bin/env python3
-
-import numpy as np
-from stdParams import *
-from src.neuronmodel import *
-from tqdm import tqdm
-
-N_sweep_distraction_scaling = 20
-N_sweep_distraction_dimension = 9
-N_samples = 1
-distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
-distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
-
-from multiprocessing import Pool
-
-
-N_p = 10
-N_out = 2
-
-T = int(2e5)
-T_test = int(1e4)
-t_ax = np.arange(T)
-
-stdMainDir = .25
-distMainDir = 2.
-
-modes = ["comp","point"]
-#modes = ["comp_bcm","point_bcm"]
-#modes = ["point"]
-
-perf = np.ndarray((N_sweep_distraction_scaling,
-                    N_sweep_distraction_dimension,
-                    N_samples,len(modes)))
-
-params = []
-
-for mode in modes:
-    for s in range(N_sweep_distraction_scaling):
-        for n in range(N_sweep_distraction_dimension):
-            for i in range(N_samples):
-                params.append([distract_dimension[n],distract_scaling[s],mode])
-
-
-
-def run_sim(arglist):
-    
-    N_distr = arglist[0]
-    s = arglist[1]
-    mode = arglist[2]
-    
-    patterns = np.random.normal(0.,1.,(T,N_p))
-    patterns[:,0] *= stdMainDir
-    patterns[:,1:N_distr+1] *= s
-    patterns[:,N_distr+1:] *= 0.
-    patterns[:,0] += (1.*(np.random.rand(T) <= 0.5) - 0.5) * distMainDir
-
-    
-    w_p = np.ones((T,N_out,N_p))
-    
-    n_d = np.ones((T,N_out))
-    n_p = np.ones((T,N_out))
-    b_d = np.zeros((T,N_out))
-    b_p = np.zeros((T,N_out))
-    
-    x_d = np.ndarray((T,2))
-    x_d[:,0] = 1.*(patterns[:,0] < 0.)
-    x_d[:,1] = 1.*(patterns[:,0] > 0.)
-    
-    #generate random orthonormal basis via qr-decomposition
-    Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
-    #Transform x_p with orthogonal basis.
[…]

### 2021-04-06 code/simulation/scaling_dimension_correlation.py

@@ -1,140 +1,0 @@
-#! /usr/bin/env python3
-
-import numpy as np
-from stdParams import *
-from src.neuronmodel import *
-from tqdm import tqdm
-
-N_sweep_distraction_scaling = 20
-N_sweep_distraction_dimension = 9
-N_samples = 2
-distract_scaling = np.linspace(1.,6.,N_sweep_distraction_scaling)
-distract_dimension = np.arange(1,N_sweep_distraction_dimension+1)
-
-align_corrcoef = np.ndarray((N_sweep_distraction_scaling,
-                    N_sweep_distraction_dimension,
-                    N_samples,2))
-
-N_p = 10
-
-T = int(5e5)
-T_test = int(1e4)
-t_ax = np.arange(T)
-
-modes = ["comp","point"]
-
-adapt_rule = "hebb"
-#adapt_rule = "fisher"
-
-for mode in tqdm(modes):
-    for s in tqdm(range(N_sweep_distraction_scaling),leave=False):
-        for n in tqdm(range(N_sweep_distraction_dimension),leave=False):
-            for i in tqdm(range(N_samples),leave=False):
-                #generate random orthonormal basis via qr-decomposition
-                Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
-                            
-                sequences = np.random.rand(T,N_p)
-                
-                w_p = np.ones((T,1,N_p))
-                
-                n_d = np.ones((T))
-                n_p = np.ones((T))
-                b_d = np.zeros((T))
-                b_p = np.zeros((T))
-                
-                x_d = (Q[:,0] @ sequences.T).T
-                
-                #Transform x_p into orthogonal basis,
-                #Scale up distract_dimension[s] dimensions
-                #and transform back.
-                x_p = (Q.T @ sequences.T).T
-                x_p[:,1:distract_dimension[n]+1] *= distract_scaling[s]
-                x_p = (Q @ x_p.T).T
-                
-                I_p = np.ndarray((T))
-                I_d = np.ndarray((T))
-                
-                x_p_av = np.ndarray((T,N_p))
-                
-                I_p_av = np.ndarray((T))
-                I_d_av = np.ndarray((T))
-                
-                y = np.ndarray((T))
-                
-                y_squ_av = np.ndarray((T))
-                y_av = np.ndarray((T))
[…]

### 2021-04-06 code/stdParams.py

@@ -18,6 +18,7 @@ thetad = 0.
 alpha = 0.3
 thetay = (1.+alpha)/2.
 mu_w = 5e-4
+eps_dec = 0.1
 mu_b = 1e-3
 mu_n = 1e-4
 mu_av = 5e-3

### 2021-04-15 notes/draft_article.tex
.

@@
 
-\usepackage{natbib}
-\bibliographystyle{apalike}
+\usepackage[numbers]{natbib}
+\bibliographystyle{abbrvnat}
 
@@
 		functional implications of the distinct physiology and anatomy of 
-		cortical pyramidal neurons. In particular, on the theoretical side,
+		cortical pyramidal neurons \cite{Spruston2008,Hay2011,Ramaswamy2015}. 
+		In particular, on the theoretical side,
 		we saw a paradigm shift from treating neurons as point-like electrical
-		structures towards embracing the entire dendritic structure. This was 
+		structures towards embracing the entire dendritic structure 
+		\cite{Larkum2009,Poirazi2009,Shai2015}. This was 
 		mostly due to the fact that experimental work uncovered dynamical properties
-		of these cells that simply could not be accounted for by point models.
+		of these cells that simply could not be accounted for by point models
+		\cite{Spruston1995,Hausser2000}.
 
@@
 		cortical pyramidal neurons can act as a separate nonlinear synaptic 
-		integration zone. Under certain conditions, a dendritic $\rm Ca^{2+}$ spike
+		integration zone \cite{Spruston2008,Branco2011}. 
+		Under certain conditions, a dendritic $\rm Ca^{2+}$ spike
 		can be elicited that propagates towards the soma, causing rapid, bursting
 		spiking activity. One of the cases in which dendritic spiking can occur
-		was termed 'backpropagation-activated $\rm Ca^{2+}$ spike firing' 
-		('BAC firing'): A single somatic spike can backpropagate towards the apical
+		was termed `backpropagation-activated $\rm Ca^{2+}$ spike firing' 
+		(`BAC firing'): A single somatic spike can backpropagate towards the apical
 		spike initiation zone, in turn significantly facilitating the initiation of 
-		a dendritic spike. This reciprocal coupling is believed to act as a form of
+		a dendritic spike \cite{Stuart2001,Spruston2008,Larkum2013}. 
+		This reciprocal coupling is believed to act as a form of
 		coincidence detection: If apical and basal synaptic input co-occurs, the 
@@
 		Naturally, these mechanisms also affect plasticity and thus learning
-		within the cortex. While the interplay between basal and apical stimulation and
+		within the cortex \cite{Sjoestroem2006,Ebner2019}. 
+		While the interplay between basal and apical stimulation and
 		its effect on synaptic efficacies is subject to ongoing research, there is
 		some evidence that BAC-firing tends to shift plasticity towards long-term potentiation
-		(LTP). Thus, coincidence between basal and apical input appears to also gate synaptic
+		(LTP) \cite{Letzkus2006}. 
+		Thus, coincidence between basal and apical input appears to also gate synaptic
 		plasticity.

### 2021-04-15 code/plotting/comp_model.py

@@ -9,16 +9,15 @@ sns.set()
 from stdParams import *
 plt.style.use('mpl_style.mplstyle')
 import os
+from src.neuronmodel import *
 
+'''
 def s(x):
     return 1./(1.+np.exp(-4.*g*x))
     
 def phi(x,y):
     return alpha * s(x)*(1.-s(y)) + s(x-thp)*s(y)
-
-alpha = 0.3
-thp = -1.
-g=1.
+'''
 
 p = np.linspace(-2.,2.,100)
 d = np.linspace(-2.,2.,100)
@@ -33,7 +32,7 @@ P,D = np.meshgrid(p,d)
 
 fig, ax = plt.subplots(1,1,figsize=(5,4))
 
-pm = ax.pcolormesh(p_ax,d_ax,phi(P,D),rasterized=True)
+pm = ax.pcolormesh(p_ax,d_ax,psi(P,D),rasterized=True)
 
 divider = make_axes_locatable(ax)
 cax = divider.append_axes("right", size="5%", pad="13%")
@@ -55,5 +54,7 @@ ax.text(2.15,-0.1,r'$\theta_{d}$',fontsize=14)
 
 fig.tight_layout(pad=0.1)
 
-fig.savefig("../../../notes/figures/plot_comp_mod_marks.pdf",transparent=True)
-fig.savefig("../../../notes/figures/plot_comp_mod_marks.png",dpi=600,transparent=True)
+fig.savefig(os.path.join(PLOT_DIR,"plot_comp_mod_marks.pdf"),transparent=True)
+fig.savefig(os.path.join(PLOT_DIR,"plot_comp_mod_marks.png"),dpi=600,transparent=True)
+
+plt.show()

### 2021-04-15 code/src/neuronmodel.py

@@ -1,12 +1,12 @@
 #! /usr/bin/env python3
 
 import numpy as np
-from stdParams import thetap0, thetap1, thetad, alpha
+from stdParams import thetap0, thetap1, thetad, alpha, gd, gp
 
 def phi(x):
     return (np.tanh(2.*x) + 1.)/2.
     
 def psi(p,d):
-    return (phi(p-thetap0)*(1.-phi(d-thetad))*alpha
-            + phi(p-thetap1)*phi(d-thetad))
+    return (phi(gp*(p-thetap0))*(1.-phi(gd*(d-thetad)))*alpha
+            + phi(gp*(p-thetap1))*phi(gd*(d-thetad)))
     

### 2021-04-15 code/stdParams.py

@@ -16,6 +16,8 @@ thetap0 = 0.
 thetap1 = -1.
 thetad = 0.
 alpha = 0.3
+gd = 1.
+gp = 1.
 thetay = (1.+alpha)/2.
 mu_w = 5e-4
 eps_dec = 0.1

### 2021-04-19 notes/draft_article.tex
.

@@
 		such as a mean square error, based on the difference between 
-		basal and apical input, i.e. $I_{\rm p} - I_{\rm d}$. Theoretical work has 
+		basal and apical input, i.e.\ $I_{\rm p} - I_{\rm d}$. Theoretical work has 
 		investigated possible learning mechanisms
@@
 		On the other hand, Hebbian-type plasticity is extensively
-		documented in experiments. Therefore, our work was
+		documented in experiments. Therefore, our work is
 		based on the question whether the non-linear interactions
@@
 
-		In our work, we combined a phenomenological model that generates the output
+		In our work, we combine a phenomenological model that generates the output
 		firing rate as a function of two streams of synaptic input (subsuming basal and apical input)
-		with Hebbian as well as BCM-like plasticity rules on basal synapses. We hypothesized that 
+		with Hebbian, as well as BCM-like plasticity rules on basal synapses. We hypothesized that 
 		this combination of neural activation and plasticity rules would lead to an
@@
 		as top-down teaching signals, without the need for an explicit error-driven
-		learning rule. Thus, we also tested our model in a simple linear 
-		supervised classification task and compared it to the performance of a simple
+		learning rule. Thus, we also test our model in a simple linear 
+		supervised classification task and compare it with the performance of a simple
 		point neuron equipped with similar plasticity rules.
@@
 
-		In all our numerical experiments, we compared this model with a 
+		In all our numerical experiments, we compare this model with a 
 		simple point neuron model that was given by
@@
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		In all our numerical experiments, the apical input $I_{\rm d}$ was generated
-		``as is", meaning, it was not dynamically calculated as a superposition
+		The apical input $I_{\rm d}$ was generated
+		`as is', meaning, it is not dynamically calculated as a superposition
 		of multiple presynaptic inputs, but given by
@@
 		discrete time sequence and $b_d(t)$ a bias. Note that $n_d$ and $b_d$ 
-		are time dependent since they were subject to adaptation processes 
+		are time dependent since they are subject to adaptation processes 
 		that are described in the next section.
 
-		Similarly, $I_{\rm p}(t)$ was given by
+		Similarly, $I_{\rm p}(t)$ is given by
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		corresponding sequences and $w_i(t)$ the synaptic efficacies.
-		As for $I_{\rm d}(t)$, $n_p(t)$ and $b_p(t)$ was used as a time dependent
+		As for $I_{\rm d}(t)$, $n_p(t)$ and $b_p(t)$ is a time dependent
 		scaling and bias.
@@
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		Additionally, we used a
+		Additionally, we use a
 		synaptic normalization constraint
@@
 
-		For comparative reasons, the point neuron model was equipped with the
+		For comparative reasons, the point neuron model is equipped with the
 		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}. 
@@
 
-		Additionally, the scaling and bias variables were changing dynamically
+		Additionally, the scaling and bias variables are changing dynamically
 		according to the following homeostatic plasticity rules:
@@
 		is shown in Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
-		basal inputs was $N=10$. One can observe a decorrelation transition for 
+		basal inputs is $N=10$. One can observe a decorrelation transition for 
 		both models. However, the compartment model supports a significantly stronger
 		distraction in terms of the scaling factor $s$ as compared to the point model.
-		This was a first confirmation of our hypothesis that nonlinear interactions
+		This is a first confirmation of our hypothesis that nonlinear interactions
 		between basal and apical input could improve learning guided by top-down signals.
@@
 		Written in vector form,
-		each sample from the basal input was generated from the following expression:
+		each sample from the basal input is generated from the following expression:
 		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		As indicated by the time dependence, the Gaussian and binary random
-		variables were drawn for each time step. The vectors
-		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ were generated
+		variables are drawn for each time step. The vectors
+		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ are generated
 		once before the beginning of a simulation run.
@@
 
-		For the classification task, we created two output neurons receiving
+		For the classification task, we set up two output neurons receiving
 		the same basal presynaptic input, while the top-down input encoded
@@
 		is shown in Fig.~\ref{fig:classification_accuracy}.
-		Albeit differences were small, the compartment model did
+		Albeit differences are small, the compartment model does
 		show a better overall accuracy for the tested parameter range.
@@
 		threshold to be the mean value between the high- and low-activity regime
-		in our compartment model, i.e.\ $\theta_M = (1+\alpha)/2$. By doing so, LTP was
-		preferably induced if both basal and apical input was present at the same
+		in our compartment model, i.e.\ $\theta_M = (1+\alpha)/2$. By doing so, LTP is
+		preferably induced if both basal and apical input are present at the same
 		time. Furthermore, instead of the weight decay term, we chose to keep
@@
 		While the accuracy of the classification for the point model
-		was at most comparable to Hebbian learning, the BCM-like rule for the 
-		compartment model significantly increased the accuracy for the tested
+		is at most comparable to Hebbian learning, the BCM-like rule for the 
+		compartment model significantly increases the accuracy for the tested
 		parameter range (compare Fig.~\ref{fig:classification_accuracy}).
 
-		Likewise, the compartment model also significantly benefited from
+		Likewise, the compartment model significantly benefits from
 		the BCM rule in terms of basal-apical alignment, while only marginal
-		improvements could be observed for the point model
+		improvements can be observed for the point model
 		(compare Fig.~\ref{fig:corr_dimension_scaling_bcm} 

### 2021-04-19 code/plotting/classification_correlation_dimension_scaling_bcm_high_input_dim.py

@@ -1,0 +1,103 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm_high_input_dim"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm_high_input_dim/"+file)))
+
+perf = []
+Ip = []
+labels = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    Ip.append(dat["I_p_samples"])
+    labels.append(dat["labels_samples"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(Ip[k].shape,Ip[k-1].shape))):
+        print("Ip arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(labels[k].shape,labels[k-1].shape))):
+        print("labels arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+Ip = np.array(Ip)
+labels = np.array(labels)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = Ip.shape[0]*Ip.shape[3]
+
+
+Ip_flatten = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],Ip.shape[5],n_total_samples))
+labels_flatten = np.ndarray((labels.shape[1],labels.shape[2],labels.shape[4],labels.shape[5],n_total_samples))
+rho = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],n_total_samples))
+
[…]

### 2021-04-19 code/plotting/classification_correlation_dimension_scaling_high_input_dim.py

@@ -1,0 +1,103 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling_high_input_dim"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling_high_input_dim/"+file)))
+
+perf = []
+Ip = []
+labels = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    Ip.append(dat["I_p_samples"])
+    labels.append(dat["labels_samples"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(Ip[k].shape,Ip[k-1].shape))):
+        print("Ip arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(labels[k].shape,labels[k-1].shape))):
+        print("labels arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+Ip = np.array(Ip)
+labels = np.array(labels)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = Ip.shape[0]*Ip.shape[3]
+
+
+Ip_flatten = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],Ip.shape[5],n_total_samples))
+labels_flatten = np.ndarray((labels.shape[1],labels.shape[2],labels.shape[4],labels.shape[5],n_total_samples))
+rho = np.ndarray((Ip.shape[1],Ip.shape[2],Ip.shape[4],n_total_samples))
+
[…]

### 2021-04-19 code/simulation/classification_dimension_scaling_bcm_high_input_dim.py

@@ -1,0 +1,198 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
+
+
+
+
+N_p = 100
+N_out = 2
+
+distract_dimension = np.linspace(1.,N_p-1,N_sweep_distraction_dimension).astype("int")
+
+T = int(2e5)
+T_test = int(1e4)
+T_sample = int(1e3)
+t_ax = np.arange(T)
+
+stdMainDir = .25
+distMainDir = 2.
+
+modes = ["comp","point"]
+#modes = ["comp_bcm","point_bcm"]
+#modes = ["point"]
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes)))
+
+I_p_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample,N_out))
+                    
+labels_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample))
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+
+def run_sim(arglist):
+    
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    patterns = np.random.normal(0.,1.,(T,N_p))
+    patterns[:,0] *= stdMainDir
+    patterns[:,1:N_distr+1] *= s
+    patterns[:,N_distr+1:] *= 0.
[…]

### 2021-04-19 code/simulation/classification_dimension_scaling_high_input_dim.py

@@ -1,0 +1,187 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
+
+
+N_p = 100
+N_out = 2
+
+distract_dimension = np.linspace(1.,N_p-1,N_sweep_distraction_dimension).astype("int")
+
+T = int(2e5)
+T_test = int(1e4)
+T_sample = int(1e3)
+t_ax = np.arange(T)
+
+stdMainDir = .25
+distMainDir = 2.
+
+modes = ["comp","point"]
+#modes = ["comp_bcm","point_bcm"]
+#modes = ["comp"]
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes)))
+                    
+I_p_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample,N_out))
+                    
+labels_samples = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes),T_sample))
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+def run_sim(arglist):
+    
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    patterns = np.random.normal(0.,1.,(T,N_p))
+    patterns[:,0] *= stdMainDir
+    patterns[:,1:N_distr+1] *= s
+    patterns[:,N_distr+1:] *= 0.
+    patterns[:,0] += (1.*(np.random.rand(T) <= 0.5) - 0.5) * distMainDir
[…]

### 2021-04-20 notes/draft_article.tex
.

@@
 		a test run without plasticity and with the apical input turned off 
-		was used to evaluate the classification	performance. For each sample,
+		was used to evaluate the classification	performance. 
+		For each sample,
 		the index of the neuron with the highest activity was used
@@
 		show a better overall accuracy for the tested parameter range.
+		It should be noted, though, that the advantage of the compartment
+		model becomes even less prominent when looking at the actual
+		correlation between proximal and distal input as a
+		measure of successful learning (as done in the previous section).
+		Fig.~\ref{fig:classification_correlation} in the supplementary 
+		section shows that the compartment model allows for a marginally
+		larger scaling among the distraction dimensions.
 
@@
 		parameter range (compare Fig.~\ref{fig:classification_accuracy}). 
+		Still, as for the Hebbian learning rule, this result should be taken
+		with a grain of salt, as Fig.~\ref{fig:classification_correlation_bcm}
+		in the supplementary section indicates that a rather large part of 
+		the region ($s>5$) showing good
+		classification performance for the compartment model in 
+		Fig.~\ref{fig:classification_accuracy_bcm} does hardly exhibit any actual
+		correlation between proximal and distal input, making it likely that
+		the accuracy of the classification could be easily impaired in this regime, 
+		e.g.\ by injecting small amounts of noise.
 
-		Likewise, the compartment model significantly benefits from
-		the BCM rule in terms of basal-apical alignment, while only marginal
+		The compartment model also significantly benefits from
+		the BCM rule in terms of basal-apical alignment as tested in Sect.~\ref{sect:alignment}, 
+		while only marginal
 		improvements can be observed for the point model
@@
 
+		\section{Supplementary Material}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{figure}
+			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling}
+			\caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning.}
+				Correlation between proximal and distal Input after training as described in 
+				Sect.~\ref{sect:classification}.}
+			\label{fig:classification_correlation}
+		\end{figure}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{figure}
+			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling_bcm}
+			\caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning
+					using a BCM-like Rule.}
+				Correlation between proximal and distal Input after linear classification 
+				training as described in Sect.~\ref{sect:classification}, but using the plasticity
+				rules given in Sect.~\ref{sect:non-hebbian}.}
+			\label{fig:classification_correlation_bcm}
+		\end{figure}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 		\bibliography{Dendritic_Learning}

### 2021-04-20 code/plotting/classification_dimension_scaling_bcm_high_input_dim.py

@@ -1,0 +1,85 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm_high_input_dim"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling_bcm_high_input_dim/"+file)))
+
+perf = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(perf[k].shape,perf[k-1].shape))):
+        print("perf arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = perf.shape[0]*perf.shape[3]
+
+perf_flatten = np.ndarray((perf.shape[1],perf.shape[2],perf.shape[4],n_total_samples))
+
+for k in range(perf.shape[0]):
+    for l in range(perf.shape[3]):
+        perf_flatten[:,:,:,k*perf.shape[3]+l] = perf[k,:,:,l,:]
+
+perf = perf_flatten
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = perf.shape[2]
+
+perf_mean = perf.mean(axis=3)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
[…]

### 2021-04-20 code/plotting/classification_dimension_scaling_high_input_dim.py

@@ -1,0 +1,85 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"classification_dimension_scaling_high_input_dim"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"classification_dimension_scaling_high_input_dim/"+file)))
+
+perf = []
+dim = []
+s = []
+
+for dat in data:
+    perf.append(dat["perf"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(perf)):
+    if(not(np.array_equal(perf[k].shape,perf[k-1].shape))):
+        print("perf arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+perf = np.array(perf)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = perf.shape[0]*perf.shape[3]
+
+perf_flatten = np.ndarray((perf.shape[1],perf.shape[2],perf.shape[4],n_total_samples))
+
+for k in range(perf.shape[0]):
+    for l in range(perf.shape[3]):
+        perf_flatten[:,:,:,k*perf.shape[3]+l] = perf[k,:,:,l,:]
+
+perf = perf_flatten
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = perf.shape[2]
+
+perf_mean = perf.mean(axis=3)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
+pc0 = ax[0].pcolormesh(s_ax,dim_ax,perf_mean[:,:,0].T,rasterized=True,vmin=0.5,vmax=1.)
[…]

### 2021-04-20 code/simulation/classification_dimension_scaling_high_input_dim.py

@@ -14,7 +14,6 @@ N_sweep_distraction_dimension = 9
 N_samples = 1
 distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
 
-
 N_p = 100
 N_out = 2
 
@@ -156,7 +155,8 @@ def run_sim(arglist):
     I_p_test = n_p[-1] * (w_p[-1] @ x_p_test.T).T - b_p[-1]
     
     pred = np.argmax(I_p_test,axis=1)
-    return (1.*(pred == lab_test)).mean(), I_p_test[:T_sample], lab_test[:T_sample]
+    
+    return (1.*(pred == lab_test)).mean(), I_p_test[:T_sample], lab_test[:T_sample], w_p, Q
 
 #'''
 pool = Pool()

### 2021-05-05 notes/draft_article.tex
.

@@
 \usepackage[numbers]{natbib}
-\bibliographystyle{abbrvnat}
+\bibliographystyle{nourlplainnat.bst}
 
@@
 		that could utilize an intracellular error signal
-		\citep{Urbanczik_2014,Schiess_2016,Guerguiev_2017}.
+		\citep{Urbanczik2014,Schiess2016,Guerguiev2017}.
 		However, a clear experimental
@@
 		On the other hand, Hebbian-type plasticity is extensively
-		documented in experiments. Therefore, our work is
+		documented in experiments \cite{Gustafsson1987,Debanne1994,Markram1997,Bi1998}. 
+		Therefore, our work is
 		based on the question whether the non-linear interactions
@@
 
-		To be continued
+		The idea of target-backpropagation in multi-layered networks
+		has already been proposed in different variants
+		\cite{Bengio2014,Lee2015,Guerguiev2017}. Yet, all of these approaches
+		assume a learning rule being dependent on an explicit error term 
+		between top-down and bottom up signals guiding plasticity in some form
+		or another.
 

### 2021-05-05 code/plotting/correlation_dimension_scaling.py

@@ -8,8 +8,6 @@ from stdParams import *
 from src.plottools import gen_mesh_ax
 plt.style.use('mpl_style.mplstyle')
 import os
-
-import os
 import sys
 
 files = os.listdir(os.path.join(DATA_DIR,"correlation_dimension_scaling"))

### 2021-05-05 code/plotting/obj_func_illustration.py

@@ -1,0 +1,37 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+from matplotlib import cm
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+from src.neuronmodel import *
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+p = np.linspace(-2.,2.,100)
+d = np.linspace(-2.,2.,100)
+
+P,D = np.meshgrid(p,d)
+
+p_ax = gen_mesh_ax(p)
+d_ax = gen_mesh_ax(d)
+
+def L(p,d):
+    return (.5*(1.-alpha)*((d-thetad)>0.)*(np.maximum(0.,p-thetap1)-np.maximum(0.,-thetap1))
+            -((alpha+1.)/2.-alpha**2.)*((d-thetad)<=0.)*np.maximum(0.,p-thetap0))
+
+fig = plt.figure()
+ax = fig.add_subplot(111, projection='3d')
+
+ax.plot_surface(P,D,L(P,D),cmap=cm.coolwarm,linewidth=0.2,rstride=6,cstride=6)
+
+ax.set_xlabel(r'$I_{\rm p}$')
+ax.set_ylabel(r'$I_{\rm d}$')
+ax.set_zlabel(r'$L$')
+
+plt.show()

### 2021-05-05 code/simulation/classification_dimension_scaling_high_input_dim.py

@@ -9,6 +9,8 @@ from datetime import datetime
 
 from multiprocessing import Pool
 
+N_processes = 6 #number of parallel processes
+
 N_sweep_distraction_scaling = 20
 N_sweep_distraction_dimension = 9
 N_samples = 1
@@ -156,14 +158,14 @@ def run_sim(arglist):
     
     pred = np.argmax(I_p_test,axis=1)
     
-    return (1.*(pred == lab_test)).mean(), I_p_test[:T_sample], lab_test[:T_sample], w_p, Q
+    return (1.*(pred == lab_test)).mean(), I_p_test[:T_sample], lab_test[:T_sample]
 
 #'''
-pool = Pool()
+pool = Pool(N_processes)
 output_list = list(tqdm(pool.imap_unordered(run_sim, params), total=len(params)))
 
-#with Pool() as p:
-#    perf_list = p.map(run_sim,params)
+#with Pool(N_processes) as p:
+#    output_list = p.map_async(run_sim,params)
 
 k = 0
 for mode in modes:

### 2021-05-05 code/simulation/correlation_dimension_scaling_bcm_high_input_dim.py

@@ -1,0 +1,187 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_processes = 6 #number of parallel processes
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
+
+N_p = 100
+N_out = 2
+
+distract_dimension = np.linspace(1.,N_p-1,N_sweep_distraction_dimension).astype("int")
+
+align_corrcoef = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,2))
+
+N_p = 10
+
+T = int(5e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+modes = ["comp","point"]
+#modes = ["point"]
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+def run_sim(arglist):
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    #generate random orthonormal basis via qr-decomposition
+    Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                
+    sequences = np.random.rand(T,N_p)
+    
+    w_p = np.ones((T,1,N_p))
+    
+    n_d = np.ones((T))
+    n_p = np.ones((T))
+    b_d = np.zeros((T))
+    b_p = np.zeros((T))
+    
+    x_d = (Q[:,0] @ sequences.T).T
+    
+    #Transform x_p into orthogonal basis,
+    #Scale up distract_dimension[s] dimensions
+    #and transform back.
+    x_p = (Q.T @ sequences.T).T
+    x_p[:,1:N_distr+1] *= s
+    x_p = (Q @ x_p.T).T
+    
[…]

### 2021-05-05 code/simulation/correlation_dimension_scaling_high_input_dim.py

@@ -1,0 +1,181 @@
+#! /usr/bin/env python3
+
+import numpy as np
+from stdParams import *
+from src.neuronmodel import *
+from tqdm import tqdm
+
+from datetime import datetime
+
+from multiprocessing import Pool
+
+N_processes = 6 #number of parallel processes
+
+N_sweep_distraction_scaling = 20
+N_sweep_distraction_dimension = 9
+N_samples = 1
+distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
+
+N_p = 100
+N_out = 2
+
+distract_dimension = np.linspace(1.,N_p-1,N_sweep_distraction_dimension).astype("int")
+
+align_corrcoef = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,2))
+
+N_p = 10
+
+T = int(5e5)
+T_test = int(1e4)
+t_ax = np.arange(T)
+
+modes = ["comp","point"]
+
+perf = np.ndarray((N_sweep_distraction_scaling,
+                    N_sweep_distraction_dimension,
+                    N_samples,len(modes)))
+
+params = []
+
+for mode in modes:
+    for s in range(N_sweep_distraction_scaling):
+        for n in range(N_sweep_distraction_dimension):
+            for i in range(N_samples):
+                params.append([distract_dimension[n],distract_scaling[s],mode])
+
+
+def run_sim(arglist):
+    N_distr = arglist[0]
+    s = arglist[1]
+    mode = arglist[2]
+    
+    #generate random orthonormal basis via qr-decomposition
+    Q,R = np.linalg.qr(np.random.normal(0.,1.,(N_p,N_p)))
+                
+    sequences = np.random.rand(T,N_p)
+    
+    w_p = np.ones((T,1,N_p))
+    
+    n_d = np.ones((T))
+    n_p = np.ones((T))
+    b_d = np.zeros((T))
+    b_p = np.zeros((T))
+    
+    x_d = (Q[:,0] @ sequences.T).T
+    
+    #Transform x_p into orthogonal basis,
+    #Scale up distract_dimension[s] dimensions
+    #and transform back.
[…]

### 2021-05-08 notes/draft_article.tex
.

@@
 		\begin{split}
-		y\left(t\right) &= \sigma\left( I_{\rm p}(t) - \theta_{p0} \right)
+		y\left(t\right) &= \alpha  \sigma\left( I_{\rm p}(t) - \theta_{p0} \right)
 		\left[1-\sigma\left(I_{\rm d}(t) - \theta_d\right)\right] \\
-		&+ \alpha \sigma\left(I_{\rm d}(t) - \theta_d \right)\sigma\left( I_{\rm p}(t) - \theta_{p1} \right)
+		&+ \sigma\left(I_{\rm d}(t) - \theta_d \right)\sigma\left( I_{\rm p}(t) - \theta_{p1} \right)
 		\end{split} \label{eq:comp_model}\\
@@
 			\left(y(t) - \tilde{y} \right)
-			\label{eq:bcm_plast} \\
+			\label{eq:hebb_plast} \\
 			\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1) \\
@@
 		For comparative reasons, the point neuron model is equipped with the
-		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}. 
+		same plasticity rule for the proximal weights as \eqref{eq:hebb_plast}. 
 %		except that the threshold $\theta_y$ was dynamically adjusted as
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{corr_dimension_scaling}
+			\includegraphics[width=\columnwidth]{corr_dimension_scaling_high_input_dim}
 			\caption{{\bf Alignment between Basal and Apical Input.} Color
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{classification_dimension_scaling}
+			\includegraphics[width=\columnwidth]{classification_dimension_scaling_high_input_dim}
 			\caption{{\bf Binary Classification Accuracy.}
@@
 		\begin{equation}
-		\Delta w_i \propto y\left(y - \theta_M\right) x_i - \epsilon w_i \; ,
+		\Delta w_i \propto y\left(y - \theta_M\right) x_i - \epsilon w_i \; , \label{eq:bcm_rule}
 		\end{equation}
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{classification_dimension_scaling_bcm}
+			\includegraphics[width=\columnwidth]{classification_dimension_scaling_bcm_high_input_dim}
 			\caption{{\bf Binary Classification Accuracy, BCM Rule.}
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{corr_dimension_scaling_bcm}
+			\includegraphics[width=\columnwidth]{corr_dimension_scaling_bcm_high_input_dim}
 			\caption{{\bf Alignment between Basal and Apical Input, BCM Rule.}
@@
 
+		\subsection{Objective Function of BCM Learning in the Compartment Model}
+		\label{sect:Obj_Func}
+
+		To form a better understanding of why the BCM-type learning rule
+		in combination with the implemented compartment model, we can formalize
+		the learning rule for the proximal weights in terms of an objective function.
+		For this purpose, we further simplify \eqref{eq:comp_model} by replacing
+		the sigmoid functions $\sigma(x)$ by a simple step function $\Theta(x)$. 
+		This does not change the overall shape or topology of the activation
+		in the $(I_p,I_d)$ space but merely makes the smooth transitions sharp
+		and instantaneous. Using $\Delta w_i \propto y\left(y - \theta_M \right) x_i$,
+		we find in this case
+		\begin{equation}
+			\begin{split}
+			\Delta w_i \propto [ (1-\alpha) \Theta(I_d - \theta_{d})\Theta(p-\theta_{p1})
+			\\ + \alpha (\alpha - 1)\Theta(\theta_{d} - I_d)\Theta(p-\theta_{p0}) ]x_i \; .
+			\end{split}
+		\end{equation}
+		Noting that $\Theta(x)$ is the first derivative of the ReLu function $[x]^+ \equiv \max(0,x)$,
+		we find that this update rule can be written as
+		\begin{align}
+			\Delta w_i &\propto \frac{\partial \mathcal{L}_p}{\partial w_i}\\
+			\begin{split}
+			\mathcal{L}_p &\equiv (1-\alpha) \Theta(I_d - \theta_{d})[p-\theta_{p1}]^+\\
+			&\;\; + \alpha (\alpha - 1)\Theta(\theta_{d} - I_d)[p-\theta_{p0}]^+ \; .
+			\end{split}\label{eq:obj_func}
+		\end{align} 
+		The objective function is shown in Fig.~\ref{fig:obj_func}. One can observe
+		that states closer to the $I_p$-$I_d$ diagonal are preferred, while the opposite is the 
+		case for off-diagonal states. This provides an explanation why the BCM-rule can induce 
+		an alignment between proximal and distal inputs when used in combination with the nonlinear 
+		compartment model. 
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+		\begin{figure}
+			\includegraphics[width=\columnwidth]{obj_func}
+			\caption{{\bf Objective Function for the Proximal Weight Update.} The 
+			approximate objective function for the proximal weights as given in
+			\eqref{eq:obj_func}. This corresponds to a combination of using 
+		\eqref{eq:comp_model} together with \eqref{eq:bcm_rule}. Note the ridge-like
+		structure along the $I_p$-$I_d$ diagonal, which supports the alignment between
+		proximal and distal input.}
+		\label{fig:obj_func}
+		\end{figure}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+		It should be noted, though, that to a certain extent, the objective function is not
+		scale-invariant in the sense that the prior distributions of both proximal 
+		and distal inputs need a certain mean and variance to cover a region of input
+		state for which the described effects can take place. This, however, is 
+		guaranteed by the homeostatic processes acting simultaneously on the gains
+		and biases.
+
 		\section{Discussion}
@@
 		between top-down and bottom up signals guiding plasticity in some form
-		or another.
+		or another. In this work, we considered an alternative approach, that is,
+		the correlation between proximal and distal input as a maximization
+		objective.
+
+		We have demonstrated that in a simple supervised learning scheme, the
+		proposed two-compartment transfer function significantly increases
+		the robustness of the learning process against distracting components
+		in the proximal input space. This was most prominently found when
+		combined with a BCM-like learning rule.
 
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling}
+			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling_high_input_dim}
 			\caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning.}
@@
 		\begin{figure}
-			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling_bcm}
+			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling_bcm_high_input_dim}
 			\caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning

### 2021-05-08 code/plotting/correlation_dimension_scaling_bcm_high_input_dim.py

@@ -1,0 +1,87 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"correlation_dimension_scaling_bcm_high_input_dim"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"correlation_dimension_scaling_bcm_high_input_dim/"+file)))
+
+rho = []
+dim = []
+s = []
+
+for dat in data:
+    rho.append(dat["align_corrcoef"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(rho)):
+    if(not(np.array_equal(rho[k].shape,rho[k-1].shape))):
+        print("corr. arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+rho = np.array(rho)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = rho.shape[0]*rho.shape[3]
+
+rho_flatten = np.ndarray((rho.shape[1],rho.shape[2],rho.shape[4],n_total_samples))
+
+for k in range(rho.shape[0]):
+    for l in range(rho.shape[3]):
+        rho_flatten[:,:,:,k*rho.shape[3]+l] = rho[k,:,:,l,:]
+
+rho = rho_flatten
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = rho.shape[2]
+
+rho_mean = rho.mean(axis=3)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
[…]

### 2021-05-08 code/plotting/correlation_dimension_scaling_high_input_dim.py

@@ -1,0 +1,85 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+files = os.listdir(os.path.join(DATA_DIR,"correlation_dimension_scaling_high_input_dim"))
+
+data = []
+
+for file in files:
+    data.append(np.load(os.path.join(DATA_DIR,"correlation_dimension_scaling_high_input_dim/"+file)))
+
+rho = []
+dim = []
+s = []
+
+for dat in data:
+    rho.append(dat["align_corrcoef"])
+    dim.append(dat["distract_dimension"])
+    s.append(dat["distract_scaling"])
+
+for k in range(1,len(rho)):
+    if(not(np.array_equal(rho[k].shape,rho[k-1].shape))):
+        print("corr. arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(dim[k],dim[k-1]))):
+        print("dim arrays do not match!")
+        sys.exit()
+    if(not(np.array_equal(s[k],s[k-1]))):
+        print("scaling arrays do not match!")
+        sys.exit()
+
+rho = np.array(rho)
+dim = dim[0]
+s = s[0]
+
+n_total_samples = rho.shape[0]*rho.shape[3]
+
+rho_flatten = np.ndarray((rho.shape[1],rho.shape[2],rho.shape[4],n_total_samples))
+
+for k in range(rho.shape[0]):
+    for l in range(rho.shape[3]):
+        rho_flatten[:,:,:,k*rho.shape[3]+l] = rho[k,:,:,l,:]
+
+rho = rho_flatten
+
+dim_ax = gen_mesh_ax(dim)
+s_ax = gen_mesh_ax(s)
+
+n_sweep = rho.shape[2]
+
+rho_mean = rho.mean(axis=3)
+
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+
[…]

### 2021-05-08 code/plotting/obj_func_illustration.py

@@ -25,13 +25,20 @@ def L(p,d):
     return (.5*(1.-alpha)*((d-thetad)>0.)*(np.maximum(0.,p-thetap1)-np.maximum(0.,-thetap1))
             -((alpha+1.)/2.-alpha**2.)*((d-thetad)<=0.)*np.maximum(0.,p-thetap0))
 
-fig = plt.figure()
+fig = plt.figure(figsize=(FIG_WIDTH,FIG_WIDTH*0.7))
 ax = fig.add_subplot(111, projection='3d')
 
 ax.plot_surface(P,D,L(P,D),cmap=cm.coolwarm,linewidth=0.2,rstride=6,cstride=6)
 
 ax.set_xlabel(r'$I_{\rm p}$')
 ax.set_ylabel(r'$I_{\rm d}$')
-ax.set_zlabel(r'$L$')
+ax.set_zlabel(r'$\mathcal{L}_{\rm p}$')
+
+ax.set_facecolor((0.,0.,0.,0.))
+
+fig.tight_layout()
+
+fig.savefig(os.path.join(PLOT_DIR,"obj_func.pdf"))
+fig.savefig(os.path.join(PLOT_DIR,"obj_func.png"),dpi=600)
 
 plt.show()

### 2021-05-09 notes/draft_article.tex
.

@@
 \bibliographystyle{nourlplainnat.bst}
+%\bibliographystyle{abbrvnat}
 
@@
 		It should be noted, though, that to a certain extent, the objective function is not
-		scale-invariant in the sense that the prior distributions of both proximal 
+		scale-invariant (as would be e.g.\ if the squared error was used) in the sense 
+		that the prior distributions of both proximal 
 		and distal inputs need a certain mean and variance to cover a region of input
-		state for which the described effects can take place. This, however, is 
-		guaranteed by the homeostatic processes acting simultaneously on the gains
+		states for which the described effects can take place. As a counterexample,
+		one could imagine that the input samples only covered a flat of $\mathcal{L}_p$, as
+		for example in Fig.~\ref{fig:obj_func} on the left, leading to a zero average gradient. 
+		This is prevented, however, by the homeostatic processes acting simultaneously on the gains
 		and biases.
@@
 		\section{Discussion}
+
+		We have demonstrated that in a simple supervised learning scheme, the
+		proposed two-compartment transfer function significantly increases
+		the robustness of the learning process against distracting components
+		in the proximal input space. This was most prominently found when
+		combined with a BCM-like learning rule.
 
@@
 		the correlation between proximal and distal input as a maximization
-		objective.
+		objective, in combination with homeostatic adaptation rules that regulate
+		proximal and distal inputs into a desired ``working regime'', 
+		see Sect.~\ref{sect:Obj_Func}. 
+		Since $I_p$ is a linear projection of the 
+		proximal input space, maximizing the correlation between
+		$I_p$ and $I_d$ can be regarded as a form of canonical correlation
+		analysis (CCA) \cite{Haerdle2007}. The idea of investigating CCA as a
+		possible mode of synaptic learning has previously been investigated by
+		\citet{Haga2018}. Interestingly, according to the authors, a 
+		BCM-learning term in the plasticity dynamics accounts for principal component
+		analysis in the input space, while CCA requires an additional multiplicative
+		term between local basal and apical activity. In contrast, our results
+		indicate that such a multiplicative term is not required to drive basal
+		synaptic plasticity towards a maximal alignment between basal and apical
+		input, even in the presence of distracting principal components.
 
-		We have demonstrated that in a simple supervised learning scheme, the
-		proposed two-compartment transfer function significantly increases
-		the robustness of the learning process against distracting components
-		in the proximal input space. This was most prominently found when
-		combined with a BCM-like learning rule.
+		As we did not include higher-dimensional distal input patterns, it
+		remains an open question how target signals would be formed in
+		multi-layered network structures. However, as previous works have
+		indicated, random top-down weights can be sufficient for
+		successful credit assignment and learning \cite{Lillicrap2016,Guerguiev2017}.
+		Therefore, our results could potentially also be transferred to deeper
+		network structures, where plasticity is guided by local errors between
+		top-down and bottom-up signals.
 

### 2021-05-09 notes/frontiers_manuscript/frontiers.tex
.

--- 2021-05-09 notes/draft_article.tex
@@
-\documentclass[10pt,a4paper,twocolumn]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
+% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.   
+%                                              %
+%                                                                                                                                                         %
+% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%% Version 3.4 Generated 2018/06/15 %%%
+%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
+%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
+%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%
+
+\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
+%\documentclass[utf8]{frontiersHLTH} % for Health articles
+%\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
+
+%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
+\usepackage{url,hyperref,lineno,microtype,subcaption}
+\usepackage[onehalfspacing]{setspace}
+\linenumbers
+
 \graphicspath{{./figures/}}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage[left=1.5cm, right=1.5cm, top=2.50cm, bottom=2.50cm]{geometry}
-\usepackage{multicol}
-\usepackage{float}
 
-\author{Fabian Schubert}
-\title{Nonlinear Dendritic Coincidence Detection for Supervised Learning}
+\def\keyFont{\fontsize{8}{11}\helveticabold }
 
-\usepackage[numbers]{natbib}
-\bibliographystyle{nourlplainnat.bst}
-%\bibliographystyle{abbrvnat}
+\def\firstAuthorLast{Fabian Schubert and Claudius Gros} %use et al 
+%only if is more than 1 author
 
-\renewcommand{\arraystretch}{1.5}
+\def\Authors{Fabian Schubert\,$^{1,*}$ and Claudius Gros\,$^{1}$}
 
-\usepackage{layouts}
+% Affiliations should be keyed to the author's name with superscript 
+%numbers and be listed as follows: Laboratory, Institute, Department, 
+%Organization, City, State abbreviation (USA, Canada, Australia), and 
+%Country (without detailed address information such as city zip codes 
+%or street names).
+% If one of the authors has a change of address, list the new address 
+%below the correspondence details using a superscript symbol and use 
+%the same symbol to indicate the author in the author list.
+
+\def\Address{$^{1}$Institute for Theoretical Physics, 
+	Goethe University Frankfurt am Main, Germany}
+
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name 
+%and city zip code) and email of the corresponding author
+
+\def\corrAuthor{Institute for Theoretical Physics\\
+	Goethe University Frankfurt am Main\\
+	Max-von-Laue-Str. 1\\
+	60438 Frankfurt am Main, Germany}
+
+\def\corrEmail{fschubert@itp.uni-frankfurt.de}
 
 \begin{document}
+\onecolumn
+\firstpage{1}
+
+\title[Running Title]{Nonlinear Dendritic Coincidence Detection for Supervised Learning}
+
+\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+\address{} %This field will be automatically populated
+\correspondance{} %This field will be automatically populated
+
+\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
+%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
+
 \maketitle
 
+\begin{abstract}
+
+Cortical pyramidal neurons have a complex dendritic anatomy, whose
+function is an active field of scientific research.
+In particular, the segregation between its soma and the apical dendritic tree
+is believed to play an active role in processing feed-forward 
+sensory information and top-down or feedback signals. In this work, we use a simple 
+two-compartment model accounting for the nonlinear interactions between
+basal and apical input streams and show that a simple Hebbian learning rule
+in the basal compartment allows the neuron to align its basal input to 
+a target signal in the apical compartment. We show that this learning
+process is robust against strong distractions in the basal input space
+and demonstrate its effectiveness in a linear classification task.
+
+\tiny
+ \keyFont{ \section{Keywords:} Dendrites, Pyramidal Neuron, Plasticity, 
+ 	Coincidence Detection, Supervised Learning}
+\end{abstract}
+
+%---------------------------------------%
 \section{Introduction}
+\label{sect:introduction}
+%---------------------------------------%
 
@@
 coincidence detection: If apical and basal synaptic input co-occurs, the 
-		neuron can respond with a rapid burst of spiking activity. The firing rate
-		of these temporal bursts exceeds the firing rate that is maximally achievable 
-		under basal synaptic input alone, therefore representing a form of temporal coincidence
+neuron can respond with a rapid burst of spiking activity. 
+The firing rate of these temporal bursts exceeds the firing 
+rate that is maximally achievable under basal synaptic input alone, 
+therefore representing a form of temporal coincidence
 detection between apical and basal input.
@@
 While the interplay between basal and apical stimulation and
-		its effect on synaptic efficacies is subject to ongoing research, there is
-		some evidence that BAC-firing tends to shift plasticity towards long-term potentiation
-		(LTP) \cite{Letzkus2006}. 
-		Thus, coincidence between basal and apical input appears to also gate synaptic
-		plasticity.
+its effect on synaptic efficacies is subject to ongoing research, 
+there is some evidence that BAC-firing tends to shift plasticity 
+towards long-term potentiation (LTP) \cite{Letzkus2006}. 
+Thus, coincidence between basal and apical input appears 
+to also gate synaptic plasticity.
 
@@
 On the other hand, Hebbian-type plasticity is extensively
-		documented in experiments \cite{Gustafsson1987,Debanne1994,Markram1997,Bi1998}. 
-		Therefore, our work is
-		based on the question whether the non-linear interactions
-		between basal and apical synaptic input could, when
-		combined with a Hebbian plasticity rule, allow a neuron
+documented in experiments 
+\cite{Gustafsson1987,Debanne1994,Markram1997,Bi1998}. 
+Therefore, our work is based on the question whether the 
+non-linear interactions between basal and apical synaptic input could, 
+when combined with a Hebbian plasticity rule, allow a neuron
 to learn to reproduce an apical teaching signal in its
@@
 
-		In our work, we combine a phenomenological model that generates the output
-		firing rate as a function of two streams of synaptic input (subsuming basal and apical input)
-		with Hebbian, as well as BCM-like plasticity rules on basal synapses. We hypothesized that 
-		this combination of neural activation and plasticity rules would lead to an
+In our work, we combine a phenomenological model that 
+generates the output firing rate as a function of two 
+streams of synaptic input (subsuming basal and apical input)
+with Hebbian, as well as BCM-like plasticity rules on basal synapses. 
+We hypothesized that this combination of neural 
+activation and plasticity rules would lead to an
 increased correlation between basal and apical input.
 
-		Furthermore, this temporal alignment could potentially facilitate apical inputs to act
-		as top-down teaching signals, without the need for an explicit error-driven
+Furthermore, this temporal alignment could potentially 
+facilitate apical inputs to act as top-down teaching signals, 
+without the need for an explicit error-driven
 learning rule. Thus, we also test our model in a simple linear 
-		supervised classification task and compare it with the performance of a simple
-		point neuron equipped with similar plasticity rules.
+supervised classification task and compare it with 
+the performance of a simple point neuron equipped with similar 
+plasticity rules.
 
+%---------------------------------------%
 \section{Model}
+\label{sect:model}
+%---------------------------------------%
 
+%---------------------------------------%
 \subsection{Neuron Model}
+\label{sect:neuronmodel}
+%---------------------------------------%
 
-		The neuron model used throughout this study is a discrete-time rate encoding model that
-		uses two separate input variables, subsuming the total synaptic
-		input current injected arriving at the basal (proximal) and apical (distal) 
-		dendritic structure of a pyramidal neuron, respectively. The model is a slightly
-		simplified version of a phenomenological model proposed by \citet{Shai_2015}.
-		Denoting the input currents by $I_{\rm p}$ (proximal) and $I_{\rm d}$ (distal), 
+The neuron model used throughout this study is a discrete-time 
+rate encoding model that uses two separate input variables, 
+subsuming the total synaptic input current injected arriving 
+at the basal (proximal) and apical (distal) 
+dendritic structure of a pyramidal neuron, respectively. 
+The model is a slightly simplified version of a phenomenological 
+model proposed by \citet{Shai_2015}. Denoting the input currents 
+$I_{\rm p}$ (proximal) and $I_{\rm d}$ (distal), 
 the model is written as
@@
 \left[1-\sigma\left(I_{\rm d}(t) - \theta_d\right)\right] \\
-		&+ \sigma\left(I_{\rm d}(t) - \theta_d \right)\sigma\left( I_{\rm p}(t) - \theta_{p1} \right)
+&+ \sigma\left(I_{\rm d}(t) - \theta_d \right)
+\sigma\left( I_{\rm p}(t) - \theta_{p1} \right)
 \end{split} \label{eq:comp_model}\\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-		Here, $\theta_{p0}$, $\theta_{p1}$ and $\theta_d$ are threshold variables with
-		respect to proximal and distal input. Overall, this equation describes two distinct
-		regions of neural activation in the $(I_{\rm p}, I_{\rm d})$-space which differ in their
-		maximal firing rates, which are set to $1$ and $\alpha$, where $0 < \alpha < 1$.
+Here, $\theta_{p0}$, $\theta_{p1}$ and $\theta_d$ are threshold 
+variables with respect to proximal and distal input. 
+Overall, this equation describes two distinct
+regions of neural activation in the 
+$(I_{\rm p}, I_{\rm d})$-space which differ in their
+maximal firing rates, which are set to $1$ and 
+$\alpha$, where $0 < \alpha < 1$.
 A plot of \eqref{eq:comp_model} is shown in Fig.~\ref{fig:comp_model}.
@@
 \centering
-			\includegraphics[width=0.8\columnwidth]{plot_comp_mod_marks.png}
+\includegraphics[width=0.6\columnwidth]{plot_comp_mod_marks.png}
 \caption{{\bf Two-compartment rate model.} Firing rate as a
@@
 
+%---------------------------------------%
 \subsection{Plasticity}
+%---------------------------------------%
 
@@
 norm of the synaptic weight vector.
-%		The threshold $\theta_y$ was
-%		chosen such that LTP is present if neural activity is in the
-%		high-activity regime (yellow area in Fig.~\ref{fig:comp_model}), while
-%		LTD is present in the low-activity regime 
-%		(given by $\alpha$, see Fig.~\ref{fig:comp_model}). Hence, we set
-%		$\theta_y = (1+\alpha)/2$.
 
@@
 same plasticity rule for the proximal weights as \eqref{eq:hebb_plast}.
-%		except that the threshold $\theta_y$ was dynamically adjusted as
-%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%		\begin{equation}
-%			\theta_y(t+1) = \left(1-\mu_{av}\right)\theta_y(t) + \mu_{av} y^2(t) \; , 
-%		\end{equation}
-%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%		which is an implementation of a running average of the square activity,
-%		as used in the classic BCM-rule.
 
@@
 b_d(t+1) &= b_d(t) + \mu_b \left[I_{\rm d}(t) - I_{\rm d}^t\right] \\
-			n_p(t+1) &= n_p(t) + \mu_n \left[V_p^t - \left( I_{\rm p}(t) - \tilde{I}_p(t)\right)^2\right] \\
-			n_d(t+1) &= n_d(t) + \mu_n \left[V_d^t - \left( I_{\rm d}(t) - \tilde{I}_d(t)\right)^2\right] \\
+n_p(t+1) &= n_p(t) + 
+\mu_n \left[V_p^t - \left( I_{\rm p}(t) - \tilde{I}_p(t)\right)^2\right] \\
+n_d(t+1) &= n_d(t) + 
+\mu_n \left[V_d^t - \left( I_{\rm d}(t) - \tilde{I}_d(t)\right)^2\right] \\
 \tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_{\rm p}(t+1) \\
@@
 Here, $I_{\rm p}^t$, $I_{\rm d}^t$, $V_p^t$ and $V_p^t$ define targets for the 
-		temporal means and variances of $I_{\rm p}$ and $I_{\rm d}$. The dynamic variables 
-		$\tilde{I}_p$ and $\tilde{I}_d$ are simply low-pass filtered running 
-		averages of $I_{\rm p}$ and $I_{\rm d}$.
+temporal means and variances of $I_{\rm p}$ and $I_{\rm d}$. 
+The dynamic variables $\tilde{I}_p$ and $\tilde{I}_d$ are simply 
+low-pass filtered running averages of $I_{\rm p}$ and $I_{\rm d}$.
 
@@
 
+%---------------------------------------%
 \section{Results}
 \label{sect:results}
+%---------------------------------------%
 
+%---------------------------------------%
 \subsection{Increased Alignment between Basal and Apical Inputs}
 \label{sect:alignment}
+%---------------------------------------%
 
@@
 A set of $N_{dist}$ orthonormal basis vectors was randomly generated,
-		which were also orthogonal to $\mathbf{a}$. Withing this $N_{dist}$-dimensional
-		subspace, the input sequences $x_{p,i}(t)$ were then rescaled by the
-		factor $s$.
+which were also orthogonal to $\mathbf{a}$. 
+Withing this $N_{dist}$-dimensional subspace, the input sequences 
+$x_{p,i}(t)$ were then rescaled by the factor $s$.
 
@@
 After this learning phase, another set of input sequences was generated
-		using the previously described protocol and $\rho[I_{\rm p},I_{\rm d}]$ was calculated.
-		Note that plasticity was turned off in this phase. This entire procedure
-		allowed us to calculate $\rho[I_{\rm p},I_{\rm d}]$ as a function of the distraction
-		parameters $s$ and $N_{dist}$. The results for both neuron models 
-		is shown in Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
+using the previously described protocol and $\rho[I_{\rm p},I_{\rm d}]$ 
+was calculated. Note that plasticity was turned off in this phase. 
+This entire procedure allowed us to calculate $\rho[I_{\rm p},I_{\rm d}]$ 
+as a function of the distraction parameters $s$ and $N_{dist}$. 
+The results for both neuron models is shown in 
+Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
 basal inputs is $N=10$. One can observe a decorrelation transition for 
-		both models. However, the compartment model supports a significantly stronger
-		distraction in terms of the scaling factor $s$ as compared to the point model.
-		This is a first confirmation of our hypothesis that nonlinear interactions
-		between basal and apical input could improve learning guided by top-down signals.
+both models. However, the compartment model supports a 
+significantly stronger distraction in terms of the scaling 
+factor $s$ as compared to the point model. This is a first confirmation 
+of our hypothesis that nonlinear interactions between basal and 
+apical input could improve learning guided by top-down signals.
 
@@
 \begin{figure}
-			\includegraphics[width=\columnwidth]{corr_dimension_scaling_high_input_dim}
+\centering
+\includegraphics[width=0.7\columnwidth]{corr_dimension_scaling_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input.} Color
@@
 
+%---------------------------------------%
 \subsection{Supervised Learning in a Linear Classification Task}
 \label{sect:classification}
+%---------------------------------------%
 
@@
 the performance in an actual supervised learning task.
-		For this purpose, we constructed presynaptic basal input as illustrated
-		in Fig.~\ref{fig:illustration_classification}.
-		Written in vector form,
-		each sample from the basal input is generated from the following expression:
+For this purpose, we constructed presynaptic basal 
+input as illustrated in Fig.~\ref{fig:illustration_classification}.
+Written in vector form, each sample from the basal 
+input is generated from the following expression:
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \centering
-			\includegraphics[width=0.75\columnwidth]{illustration_classification}
+\includegraphics[width=0.55\columnwidth]{illustration_classification}
 \caption{{\bf Input Space for a Simple Linear Classification Task.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 For the classification task, we set up two output neurons receiving
@@
 \begin{align}
-			x_{d,0}(t) &= 1 - \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \\
-			x_{d,1}(t) &= \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \; ,
+x_{d,0}(t) &= 1 - \Theta\left( \left(\mathbf{x}_p(t) - 
+\mathbf{b}\right)^T \mathbf{a}\right) \\
+x_{d,1}(t) &= \Theta\left( \left(\mathbf{x}_p(t) - 
+\mathbf{b}\right)^T \mathbf{a}\right) \; ,
 \end{align}
@@
 was used to evaluate the classification	performance. 
-		For each sample,
-		the index of the neuron with the highest activity was used
-		as the predicted class. Accuracy was then calculated as the fraction
+For each sample, the index of the neuron with the highest 
+activity was used as the predicted class. 
+Accuracy was then calculated as the fraction
 of correctly classified samples.
@@
 \begin{figure}
-			\includegraphics[width=\columnwidth]{classification_dimension_scaling_high_input_dim}
+\centering
+\includegraphics[width=0.7\columnwidth]{classification_dimension_scaling_high_input_dim}
 \caption{{\bf Binary Classification Accuracy.}
@@
 larger scaling among the distraction dimensions.
-
+\\
+%---------------------------------------%
 \subsection{Non-Hebbian Learning Rules}
 \label{sect:non-hebbian}
+%---------------------------------------%
 
@@
 \begin{equation}
-		\Delta w_i \propto y\left(y - \theta_M\right) x_i - \epsilon w_i \; , \label{eq:bcm_rule}
+\Delta w_i \propto y\left(y - \theta_M\right) x_i - 
+\epsilon w_i \; , \label{eq:bcm_rule}
 \end{equation}
@@
 The compartment model also significantly benefits from
-		the BCM rule in terms of basal-apical alignment as tested in Sect.~\ref{sect:alignment}, 
-		while only marginal
+the BCM rule in terms of basal-apical alignment as tested in 
+Sect.~\ref{sect:alignment}, while only marginal
 improvements can be observed for the point model
@@
 with Fig.~\ref{fig:corr_dimension_scaling}).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
-			\includegraphics[width=\columnwidth]{classification_dimension_scaling_bcm_high_input_dim}
+\centering
+\includegraphics[width=0.7\columnwidth]{classification_dimension_scaling_bcm_high_input_dim}
 \caption{{\bf Binary Classification Accuracy, BCM Rule.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
-			\includegraphics[width=\columnwidth]{corr_dimension_scaling_bcm_high_input_dim}
+\centering
+\includegraphics[width=0.7\columnwidth]{corr_dimension_scaling_bcm_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input, BCM Rule.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\\
+\\
+%---------------------------------------%
 \subsection{Objective Function of BCM Learning in the Compartment Model}
 \label{sect:Obj_Func}
+%---------------------------------------%
 
@@
 we find in this case
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{equation}
@@
 \end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Noting that $\Theta(x)$ is the first derivative of the ReLu function $[x]^+ \equiv \max(0,x)$,
 we find that this update rule can be written as
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{align}
@@
 \end{align}
-		The objective function is shown in Fig.~\ref{fig:obj_func}. One can observe
-		that states closer to the $I_p$-$I_d$ diagonal are preferred, while the opposite is the 
-		case for off-diagonal states. This provides an explanation why the BCM-rule can induce 
-		an alignment between proximal and distal inputs when used in combination with the nonlinear 
-		compartment model. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The objective function is shown in Fig.~\ref{fig:obj_func}. 
+One can observe that states closer to the $I_p$-$I_d$ diagonal 
+are preferred, while the opposite is the case for off-diagonal states. 
+This provides an explanation why the BCM-rule can induce 
+an alignment between proximal and distal inputs when used 
+in combination with the nonlinear compartment model. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
-			\includegraphics[width=\columnwidth]{obj_func}
+\centering
+\includegraphics[width=0.7\columnwidth]{obj_func}
 \caption{{\bf Objective Function for the Proximal Weight Update.} The 
@@
 
-		It should be noted, though, that to a certain extent, the objective function is not
-		scale-invariant (as would be e.g.\ if the squared error was used) in the sense 
+It should be noted, though, that to a certain extent, the objective 
+function is not scale-invariant (as would be e.g.\ if 
+the squared error was used) in the sense 
 that the prior distributions of both proximal 
 and distal inputs need a certain mean and variance to cover a region of input
-		states for which the described effects can take place. As a counterexample,
-		one could imagine that the input samples only covered a flat of $\mathcal{L}_p$, as
-		for example in Fig.~\ref{fig:obj_func} on the left, leading to a zero average gradient. 
-		This is prevented, however, by the homeostatic processes acting simultaneously on the gains
+states for which the described effects can take place. 
+As a counterexample, one could imagine that the input samples 
+only covered a flat of $\mathcal{L}_p$, as for example in 
+Fig.~\ref{fig:obj_func} on the left, leading to a zero average gradient. 
+This is prevented, however, by the homeostatic processes 
+acting simultaneously on the gains
 and biases.
 
+%---------------------------------------%
 \section{Discussion}
+%---------------------------------------%
 
@@
 
+%---------------------------------------%
 \section{Supplementary Material}
+%---------------------------------------%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}
-			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling_high_input_dim}
+\centering
+\includegraphics[width=0.7\columnwidth]{classification_correlation_dimension_scaling_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning.}
@@
 \begin{figure}
-			\includegraphics[width=\columnwidth]{classification_correlation_dimension_scaling_bcm_high_input_dim}
+\centering
+\includegraphics[width=0.7\columnwidth]{classification_correlation_dimension_scaling_bcm_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning
@@
 
-		\bibliography{Dendritic_Learning}
+%\section{Manuscript Formatting}
+%
+%\subsection{Heading Levels}
+%
+%%There are 5 heading levels
+%
+%\subsection{Level 2}
+%\subsubsection{Level 3}
+%\paragraph{Level 4}
+%\subparagraph{Level 5}
+%
+%\subsection{Equations}
+%Equations should be inserted in editable format from the equation editor.
+%
+%\begin{equation}
+%\sum x+ y =Z\label{eq:01}
+%\end{equation}
+%
+%\subsection{Figures}
+%Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
+Figures will then be automatically embedded at the bottom of the submitted manuscript.
+Kindly ensure that each table and figure is mentioned in the text and in numerical order.
+Figures must be of sufficient resolution for publication \href{http://home.frontiersin.org/about/author-guidelines#ResolutionRequirements}{see here for examples and minimum requirements}.
+Figures which are not according to the guidelines will cause substantial delay during the production process.
+Please see \href{http://home.frontiersin.org/about/author-guidelines#GeneralStyleGuidelinesforFigures}{here} for full figure guidelines.
+Cite figures with subfigures as figure \ref{fig:2}B.
+%
+%
+%\subsubsection{Permission to Reuse and Copyright}
+%Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
+It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
+%%%Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
+It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
+%
+%\subsection{Tables}
+%Tables should be inserted at the end of the manuscript.
+Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
+Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
+These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
+The author will be notified during the typesetting of the final article if this is the case.
+%
+%\section{Nomenclature}
+%
+%\subsection{Resource Identification Initiative}
+%To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript.
+For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
+%
+%\subsection{Life Science Identifiers}
+%Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords.
+For more information on LSIDs please see \href{http://www.frontiersin.org/about/AuthorGuidelines#InclusionofZoologicalNomenclature}{Inclusion of Zoological Nomenclature} section of the guidelines.
+%
+%
+%\section{Additional Requirements}
+%
+%For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.
+
+\section*{Conflict of Interest Statement}
+%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
+If no such relationship exists, authors will be asked to confirm the following statement:
+
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+\section*{Author Contributions}
+
+Both authors, F.S. and C.G., contributed equally to the
+writing and review of the manuscript. F.S. provided the code,
+ran the simulations and prepared the figures.
+
+%\section*{Funding}
+%Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+
+%---------------------------------------%
+\section*{Acknowledgments}
+%---------------------------------------%
+
+The authors acknowledge the financial support of
+the German research foundation (DFG)
+\section*{Supplemental Data}
+ \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} 
+ should be uploaded separately on submission, if there are Supplementary Figures, 
+ please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
+
+\section*{Data Availability Statement}
+The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
+% Please see the availability of data guidelines for more information, 
+%at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
+
+\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities 
+%and Social Sciences articles, for Humanities and Social 
+%Sciences articles please include page numbers in the in-text citations
+%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, 
+%Physics and Mathematics articles
+\bibliography{Dendritic_Learning.bib}
+
+%%% Make sure to upload the bib file along with the tex file and PDF
+%%% Please see the test.bib file for some examples of references
+
+%\section*{Figure captions}
+%
+%%%% Please be aware that for original research articles we only permit 
+%%a combined number of 15 figures and tables, one figure with multiple 
+%%subfigures will count as only one figure.
+%%%% Use this if adding the figures directly in the mansucript, if so, 
+%%please remember to also upload the files when submitting your article
+%%%% There is no need for adding the file termination, as long as 
+%%you indicate where the file is saved. In the examples below the 
+%%files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
+%%%% If using *.tif files convert them to .jpg or .png
+%%%%  NB logo1.eps is required in the path in order to correctly 
+%%compile front page header %%%
+%
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[width=10cm]{logo1}% This is a *.eps file
+%\end{center}
+%\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
+%\end{figure}
+%
+%
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[width=15cm]{logos}
+%\end{center}
+%\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
+%\end{figure}
+%
+%%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
+%%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
+%%%% Frontiers will add the figures at the end of the provisional pdf automatically
+%%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.
 

### 2021-05-12 notes/frontiers_coincidence_learning/frontiers.tex
.

--- 2021-05-09 notes/frontiers_manuscript/frontiers.tex
@@
 \def\keyFont{\fontsize{8}{11}\helveticabold }
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Cortical pyramidal neurons have a complex dendritic anatomy, whose
-function is an active field of scientific research.
-In particular, the segregation between its soma and the apical dendritic tree
-is believed to play an active role in processing feed-forward 
-sensory information and top-down or feedback signals. In this work, we use a simple 
-two-compartment model accounting for the nonlinear interactions between
-basal and apical input streams and show that a simple Hebbian learning rule
-in the basal compartment allows the neuron to align its basal input to 
-a target signal in the apical compartment. We show that this learning
-process is robust against strong distractions in the basal input space
-and demonstrate its effectiveness in a linear classification task.
+Cortical pyramidal neurons have a complex 
+dendritic anatomy, whose function is an active 
+field of scientific research. In particular, 
+the segregation between its soma and the apical 
+dendritic tree is believed to play an active 
+role in processing feed-forward sensory 
+information and top-down or feedback signals. 
+In this work, we use a simple two-compartment 
+model accounting for the nonlinear interactions 
+between basal and apical input streams and show 
+that a simple Hebbian learning rule in the basal 
+compartment allows the neuron to align its basal 
+input to a target signal in the apical compartment. 
+We show that this learning process, termed 
+coincidence detection, is robust against strong 
+distractions in the basal input space and 
+demonstrate its effectiveness in a linear 
+classification task.
 
@@
 functional implications of the distinct physiology and anatomy of 
-cortical pyramidal neurons \cite{Spruston2008,Hay2011,Ramaswamy2015}. 
+cortical pyramidal neurons \citep{Spruston2008,Hay2011,Ramaswamy2015}. 
 In particular, on the theoretical side,
@@
 structures towards embracing the entire dendritic structure 
-\cite{Larkum2009,Poirazi2009,Shai2015}. This was 
-mostly due to the fact that experimental work uncovered dynamical properties
-of these cells that simply could not be accounted for by point models
-\cite{Spruston1995,Hausser2000}.
+\citep{Larkum2009,Poirazi2009,Shai2015}. This was 
+mostly due to the fact that experimental work 
+uncovered dynamical properties of pyramidal neuronal
+cells that simply could not be accounted for by point models
+\citep{Spruston1995,Hausser2000}.
 
-An important finding was that the apical dendritic tree of
+An important finding is that the apical dendritic tree of
 cortical pyramidal neurons can act as a separate nonlinear synaptic 
-integration zone \cite{Spruston2008,Branco2011}. 
+integration zone \citep{Spruston2008,Branco2011}. 
 Under certain conditions, a dendritic $\rm Ca^{2+}$ spike
@@
 spike initiation zone, in turn significantly facilitating the initiation of 
-a dendritic spike \cite{Stuart2001,Spruston2008,Larkum2013}. 
+a dendritic spike \citep{Stuart2001,Spruston2008,Larkum2013}. 
 This reciprocal coupling is believed to act as a form of
@@
 Naturally, these mechanisms also affect plasticity and thus learning
-within the cortex \cite{Sjoestroem2006,Ebner2019}. 
+within the cortex \citep{Sjoestroem2006,Ebner2019}. 
 While the interplay between basal and apical stimulation and
 its effect on synaptic efficacies is subject to ongoing research, 
-there is some evidence that BAC-firing tends to shift plasticity 
-towards long-term potentiation (LTP) \cite{Letzkus2006}. 
+there is evidence that BAC-firing tends to shift plasticity 
+towards long-term potentiation (LTP) \citep{Letzkus2006}. 
 Thus, coincidence between basal and apical input appears 
@@
 such as a mean square error, based on the difference between 
-basal and apical input, i.e.\ $I_{\rm p} - I_{\rm d}$. Theoretical work has 
-investigated possible learning mechanisms
-that could utilize an intracellular error signal
+basal and apical input, i.e.\ $I_p - I_d$,
+where indices $p$ and $d$ denote `proximal' and
+`distal', in equivalence to basal and apical. 
+Theoretical studies have investigated possible 
+learning mechanisms that could utilize an 
+intracellular error signal
 \citep{Urbanczik2014,Schiess2016,Guerguiev2017}.
@@
 documented in experiments 
-\cite{Gustafsson1987,Debanne1994,Markram1997,Bi1998}. 
+\citep{Gustafsson1987,Debanne1994,Markram1997,Bi1998}. 
 Therefore, our work is based on the question whether the 
@@
 
-In our work, we combine a phenomenological model that 
-generates the output firing rate as a function of two 
-streams of synaptic input (subsuming basal and apical input)
-with Hebbian, as well as BCM-like plasticity rules on basal synapses. 
-We hypothesized that this combination of neural 
+We investigate coincidence learning by
+combining a phenomenological model that 
+generates the output firing rate as a function 
+of two streams of synaptic input (subsuming basal 
+and apical inputs) with Hebbian, as well as 
+BCM-like plasticity rules on basal synapses. 
+In particular we hypothesized that this combination of neural 
 activation and plasticity rules would lead to an
-increased correlation between basal and apical input.
-
-Furthermore, this temporal alignment could potentially 
-facilitate apical inputs to act as top-down teaching signals, 
-without the need for an explicit error-driven
-learning rule. Thus, we also test our model in a simple linear 
-supervised classification task and compare it with 
-the performance of a simple point neuron equipped with similar 
-plasticity rules.
+increased correlation between basal and apical inputs.
+Furthermore, the temporal alignment observed in our study 
+could potentially facilitate apical inputs to act as 
+top-down teaching signals, without the need for an 
+explicit error-driven learning rule. Thus, we also 
+test our model in a simple linear supervised 
+classification task and compare it with the 
+performance of a simple point neuron equipped with 
+similar plasticity rules.
 
@@
 
-The neuron model used throughout this study is a discrete-time 
-rate encoding model that uses two separate input variables, 
+The neuron model used throughout this study 
+is a discrete-time rate encoding model that 
+contains two separate input variables, 
 subsuming the total synaptic input current injected arriving 
@@
 model proposed by \citet{Shai_2015}. Denoting the input currents 
-$I_{\rm p}$ (proximal) and $I_{\rm d}$ (distal), 
+$I_p$ (proximal) and $I_d$ (distal), 
 the model is written as
@@
 \begin{split}
-y\left(t\right) &= \alpha  \sigma\left( I_{\rm p}(t) - \theta_{p0} \right)
-\left[1-\sigma\left(I_{\rm d}(t) - \theta_d\right)\right] \\
-&+ \sigma\left(I_{\rm d}(t) - \theta_d \right)
-\sigma\left( I_{\rm p}(t) - \theta_{p1} \right)
-\end{split} \label{eq:comp_model}\\
+y\left(t\right) &= \alpha  \sigma\left( I_p(t) - \theta_{p0} \right)
+\left[1-\sigma\left(I_d(t) - \theta_d\right)\right] \\
+&+ \sigma\left(I_d(t) - \theta_d \right)
+\sigma\left( I_p(t) - \theta_{p1} \right)
+\end{split} 
+\label{eq_comp_model}\\
 \sigma(x) &\equiv \frac{1}{1+\exp(-4x)} \; .
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Here, $\theta_{p0}$, $\theta_{p1}$ and $\theta_d$ are threshold 
-variables with respect to proximal and distal input. 
-Overall, this equation describes two distinct
+Here, $\theta_{p0}>\theta_{p1}$ and $\theta_d$ 
+are threshold variables with respect to proximal 
+and distal inputs. Overall, equation 
+(\ref{eq_comp_model}) describes two distinct
 regions of neural activation in the 
-$(I_{\rm p}, I_{\rm d})$-space which differ in their
+$(I_p, I_d)$-space which differ in their
 maximal firing rates, which are set to $1$ and 
 $\alpha$, where $0 < \alpha < 1$.
-A plot of \eqref{eq:comp_model} is shown in Fig.~\ref{fig:comp_model}.
+A plot of \eqref{eq_comp_model} is shown 
+in Fig.~\ref{fig:comp_model}.
+
+When both input currents $I_d$ and $I_p$ 
+are large, viz larger than the 
+thresholds $\theta_d$ and $\theta_{p1}$,
+the second term in \eqref{eq_comp_model}
+dominates, which leads to $y\approx 1$. 
+An intermediate activity plateau, of
+strength $\alpha$ emerges in addition 
+when $I_p>\theta_{p0}$ and 
+$I_d<\theta_{d}$. As such, the compartment
+model \eqref{eq_comp_model} is able to
+distinguish neurons with a normal activity 
+level, here encoded by $\alpha=0.3$, and
+strongly bursting neurons, where the maximal
+firing rate is unity. The intermediate plateau
+allows neurons to process the proximal 
+inputs $I_p$ even in the absence of distal
+stimulation. The distal current $I_d$
+acts therefore as an additional modulator.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
 \includegraphics[width=0.6\columnwidth]{plot_comp_mod_marks.png}
-\caption{{\bf Two-compartment rate model.} Firing rate as a
-	function of proximal and distal input $I_{\rm p}$ and $I_{\rm d}$, 
-	see \eqref{eq:comp_model}. 
-	The thresholds $\theta_{p0}$, $\theta_{p1}$ and $\theta_d$ define
-	two regions of neural activity with a a maximal firing rate of $1$
-	and $\alpha$.}
+\caption{{\bf Two-compartment rate model.} 
+The firing rate as a function of proximal and distal 
+inputs $I_p$ and $I_d$, 
+see \eqref{eq_comp_model}. 
+The thresholds $\theta_{p0}$, $\theta_{p1}$ and 
+$\theta_d$ define two regions of neural activity,
+with a a maximal firing rate of unity an a plateau
+at $\alpha=0.3$.}
 \label{fig:comp_model}
@@
 
-In all our numerical experiments, we compare this model with a 
-simple point neuron model that was given by
+In our numerical experiments we compare 
+the compartment model with a classical point 
+neuron, as given by
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{equation}
-y(t) = \sigma\left(I_{\rm p}(t) + I_{\rm d}(t) - \theta \right) \; .
+y(t) = \sigma\left(I_p(t) + I_d(t) - 
+\theta \right) \; .
+\label{eq_point_neuron}
 \end{equation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The apical input $I_{\rm d}$ was generated
-`as is', meaning, it is not dynamically calculated as a superposition
-of multiple presynaptic inputs, but given by
+The apical input $I_d$ is generated
+`as is', meaning, it is not dynamically 
+calculated as a superposition of multiple 
+presynaptic inputs. To be concrete, we
+used
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{equation}
-I_{\rm d}(t) = n_d(t) x_d(t) - b_d(t) \; ,
-\label{eq:I_d}
+I_d(t) = n_d(t) x_d(t) - b_d(t) \; ,
+\label{eq_I_d}
 \end{equation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $n_d(t)$ is a scaling factor, $x_d(t)$ a pre-generated
-discrete time sequence and $b_d(t)$ a bias. Note that $n_d$ and $b_d$ 
-are time dependent since they are subject to adaptation processes 
-that are described in the next section.
-
-Similarly, $I_{\rm p}(t)$ is given by
+where $n_d(t)$ is a scaling factor, $x_d(t)$ 
+a pre-generated discrete time sequence and 
+$b_d(t)$ a bias. Note that $n_d$ and $b_d$ 
+are time dependent since they are subject 
+to adaptation processes, which will be
+described in the next section. Similarly, 
+the proximal input $I_p(t)$ is given by
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{equation}
-I_{\rm p}(t) = n_p(t) \sum_{i=1}^{N} x_{p,i}(t) w_i(t) - b_p(t) \; ,
+I_p(t) = n_p(t) \sum_{i=1}^{N} 
+x_{p,i}(t) w_i(t) - b_p(t) \; ,
+\label{eq_I_p}
 \end{equation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $N$ is the number of presynaptic afferents, $x_{p,i}(t)$ the
-corresponding sequences and $w_i(t)$ the synaptic efficacies.
-As for $I_{\rm d}(t)$, $n_p(t)$ and $b_p(t)$ is a time dependent
-scaling and bias.
+where $N$ is the number of presynaptic afferents, 
+$x_{p,i}(t)$ the corresponding sequences, 
+$w_i(t)$ the synaptic efficacies and
+$n_p(t)$ and $b_p(t)$ the (time dependent)
+scaling and bias. Tyical values for the 
+parameters used throughout this study
+are presented Table~\ref{tab_parameters}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[b]
+\caption{Model parameters,
+as defined in sections \ref{sect:neuronmodel} and
+\ref{sect_plasticity}.}
+\begin{tabular}{ l | l || l | l }
+		$\theta_{p0}$ & $0$ & $V_d^t$ & $0.25$ \\
+		$\theta_{p1}$ & $-1$ & $\mu_b$ & $10^{-3}$ \\ 
+		$\theta_{d}$ & $0$ & $\mu_n$ & $10^{-4}$ \\  
+		$\alpha$ & $0.3$ & $\mu_{\rm av}$ & $5 \cdot 10^{-3}$ \\   
+		$\mu_w$ & $5 \cdot 10^{-5}\quad$ & $I_p^t$ & $0$ \\
+		$V_p^t$ & $0.25$ & $I_d^t$ & $0$  
+\end{tabular}
+\label{tab_parameters}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 %---------------------------------------%
-\subsection{Plasticity}
+\subsection{Plasticity\label{sect_plasticity}}
 %---------------------------------------%
 
-We implemented a Hebbian plasticity rule for the basal 
-synaptic weights given by the following update equation:
+We implemented a Hebbian plasticity rule for 
+the proximal synaptic weights by the following 
+update equation:
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \left(y(t) - \tilde{y} \right)
-\label{eq:hebb_plast} \\
-\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1) \\
-\tilde{y}(t+1) &= (1-\mu_{av})\tilde{y}(t) + \mu_{av}y(t+1)
+\label{eq_hebb_plast} \\
+\tilde{x}_{p,i}(t+1) &= 
+(1-\mu_{\rm av})\tilde{x}_{p,i}(t) + \mu_{\rm av}x_{p,i}(t) 
+\label{eq_tilde_x_p_i}\\
+\tilde{y}(t+1) &= (1-\mu_{\rm av})\tilde{y}(t) + \mu_{\rm av}y(t)
+\label{eq_tilde_y}
 \end{align}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Additionally, we use a
-synaptic normalization constraint
+The trailing time averages $\tilde{x}_{p,i}$ and
+$\tilde{y}$, respectively of the presynaptic 
+basal activites, $x_{p,i}$, and of the neural
+firing rate $y$, enter the Hebbian learing
+rule (\ref{eq_hebb_plast}) as reference levels.
+Pre- and post-synaptic neurons are considered
+to be active/inactive when being above/below
+the respective trailing averages. The
+timescale of the averaging, $1/\mu_{\rm av}$,
+is typcially over 200 time steps, see
+Table~\ref{tab_parameters}. 
+With $1/\mu_w=2\cdot10^4$ learning is 
+assumed to be be considerably slower, 
+as usual for statistical update rules.
+Additionally, 
+we use a synaptic normalization constraint,
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{equation}
-w_i(t) \rightarrow \frac{w_i(t)}{\lVert \mathbf{w}(t)\rVert}
+w_i(t) \rightarrow \frac{w_i(t)}{\lVert 
+\mathbf{w}(t)\rVert}\;,
 \label{eq:weight_norm}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-in each time step, where $\lVert \mathbf{w}(t)\rVert$ denotes the Euclidean
-norm of the synaptic weight vector.
+in each time step, where 
+$\lVert \mathbf{w}(t)\rVert$ denotes the Euclidean
+norm of the synaptic weight vector. For comparative 
+reasons, the point neuron model is equipped with the
+same plasticity rule for the proximal weights as 
+(\ref{eq_hebb_plast}).
 
-For comparative reasons, the point neuron model is equipped with the
-same plasticity rule for the proximal weights as \eqref{eq:hebb_plast}.
-
-Additionally, the scaling and bias variables are changing dynamically
-according to the following homeostatic plasticity rules:
+The bias variables entering the definitions
+\eqref{eq_I_d} and \eqref{eq_I_p} of the distal
+proximal currrent, $I_d$ and $I_p$, 
+are assumed to adapt accodingly to
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{align}
-b_p(t+1) &= b_p(t) + \mu_b \left[I_{\rm p}(t) - I_{\rm p}^t\right] \\
-b_d(t+1) &= b_d(t) + \mu_b \left[I_{\rm d}(t) - I_{\rm d}^t\right] \\
-n_p(t+1) &= n_p(t) + 
-\mu_n \left[V_p^t - \left( I_{\rm p}(t) - \tilde{I}_p(t)\right)^2\right] \\
-n_d(t+1) &= n_d(t) + 
-\mu_n \left[V_d^t - \left( I_{\rm d}(t) - \tilde{I}_d(t)\right)^2\right] \\
-\tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_{\rm p}(t+1) \\
-\tilde{I}_d(t+1) &= (1-\mu_{av})\tilde{I}_d(t) + \mu_{av} I_{\rm d}(t+1)
+\label{eq_b_p_dot}
+b_p(t+1) &= b_p(t) + \mu_b \left[I_p(t) - I_p^t\right] \\
+b_d(t+1) &= b_d(t) + \mu_b \left[I_d(t) - I_d^t\right] \;,
+\label{eq_b_d_dot}
 \end{align}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Here, $I_{\rm p}^t$, $I_{\rm d}^t$, $V_p^t$ and $V_p^t$ define targets for the 
-temporal means and variances of $I_{\rm p}$ and $I_{\rm d}$. 
-The dynamic variables $\tilde{I}_p$ and $\tilde{I}_d$ are simply 
-low-pass filtered running averages of $I_{\rm p}$ and $I_{\rm d}$.
+where $I_p^t=0$ and, $I_d^t=0$
+are preset targets and $1/\mu_b=10^3$ the
+timescale for the adaption process. Over
+time, both the distal and the proximal 
+currents, $I_d$ and $I_p$, average out.
 
-A list of all parameter values is given in Table~\ref{tab:parameters}.
+Adaption rules for the bias entering a 
+transfer function, such as
+\eqref{eq_b_d_dot} and \eqref{eq_b_p_dot},
+have the task to regulate overall activitiy
+levels. The overall magnitude of the
+synaptic weights, which are determined by 
+synaptic rescaling factors, here $n_d$ and $n_p$, 
+as defined in \eqref{eq_I_d} and \eqref{eq_I_p},
+will regulate in contrast the variance of
+the neural activity, and not the average
+level \citep{schubert2021local}. In this
+spirit we consider
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{align}
+n_d(t+1) &= n_d(t) + 
+\mu_n \left[V_d^t - \left( I_d(t) - \tilde{I}_d(t)\right)^2\right] \\
+n_p(t+1) &= n_p(t) + 
+\mu_n \left[V_p^t - \left( I_p(t) - \tilde{I}_p(t)\right)^2\right] \\
+\tilde{I}_d(t+1) &= (1-\mu_{\rm av})\tilde{I}_d(t) + \mu_{\rm av} I_d(t) \\
+\tilde{I}_p(t+1) &= (1-\mu_{\rm av})\tilde{I}_p(t) + \mu_{\rm av} I_p(t) 
+\end{align}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Here, $V_p^t$ and $V_p^t$ define targets for 
+the temporal averaged variances of $I_p$ 
+and $I_d$.  
+The dynamic variables $\tilde{I}_p$ and $\tilde{I}_d$ 
+are simply low-pass filtered running averages of 
+$I_p$ and $I_d$. Overall, the framework
+specified here allows the neuron to be fully flexible,
+as long as the activtiy level and its variance 
+fluctuate around preset target values 
+\citep{schubert2021local}. A list of 
+the parameter values used throughout this investigation 
+is given in Table~\ref{tab_parameters}. Our choices of
+target means and variances are based on the assumption
+that neural input should be tuned towards a certain
+working regime of the neural transfer function. In
+the case of the presented model, this means that both
+proximal and distal input cover an area where the nonlinearities
+of the transfer function are reflected without oversaturation.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}
-	\begin{tabular}{ l | l || l | l }
-		$\theta_{p0}$ & $0$ & $V_d^t$ & $0.25$ \\
-		$\theta_{p1}$ & $-1$ & $\mu_b$ & $10^{-3}$ \\ 
-		$\theta_{d}$ & $0$ & $\mu_n$ & $10^{-4}$ \\  
-		$\alpha$ & $0.3$ & $\mu_{av}$ & $5 \cdot 10^{-3}$ \\   
-		$\mu_w$ & $5 \cdot 10^{-5}$ & $I_{\rm p}^t$ & $0$ \\
-		$V_p^t$ & $0.25$ & $I_{\rm d}^t$ & $0$  
-	\end{tabular}
-	\caption{Model parameters}
-	\label{tab:parameters}
-\end{table}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.55\columnwidth]{illustration_classification}
+\caption{{\bf Input Space for the Linear Classification Task.}
+Two clusters of presynaptic basal activities were generated from 
+multivariate Gaussian distributions. Here, $s$ denotes the standard
+deviation orthogonal to the normal vector $\mathbf{a}$ of the 
+classification hyperplane, as defined by \eqref{eq_x_d_a}.}
+\label{fig_illustration_classification}
+\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %---------------------------------------%
-\subsection{Increased Alignment between Basal and Apical Inputs}
+\subsection{Alignment between Basal and Apical Inputs}
 \label{sect:alignment}
@@
 
-As a first test, we wanted to quantify the neuron's ability to 
+As a first test, we quantify the neuron's ability to 
 align its basal input to the apical teaching signal.
-To do so, we defined the pearson correlation coefficient
-$\rho[I_{\rm p},I_{\rm d}]$ between the basal and apical input current
-as our measure of interest. We determined this temporal
-correlation coefficient after simulating all plasticity mechanisms
-under stationary random input sequences with certain statistical
-properties that shall be explained in the following.
-As a starting point, we chose all $x_{p,i}(t)$ to be randomly
-drawn from a uniform distribution, where $x_{p,i}(t) \in (0,1)$.
-For $I_{\rm d}(t)$ to be fully `reconstructable' by the basal input, 
-$x_d(t)$ had to be some linear combination of all $x_{p,i}(t)$.
-Therefore, we chose $x_d(t) = \sum_{i=1}^N a_i x_{p,i}(t)$, where
-$\mathbf{a}$ is a random vector with unit length.
-Since we used a Hebbian learning scheme, we expected that 
-the direction and magnitude of the principal components of 
-the basal input would also significantly affect the outcome of
-the experiment: A large variance in the basal input 
+This can be done using the pearson correlation coefficient
+$\rho[I_p,I_d]$ between the basal and 
+apical input currents. We determined 
+$\rho[I_p,I_d]$ after the simulation,
+which involves all plasticity mechanisms, both
+for the synaptic weights and for the intrinsic 
+parameters. The input sequences $x_{p,i}(t)$ 
+is randomly drawn from a uniform distribution,
+in $[0,1]$, which is done independently for 
+each $i\in[1,N]$.
+
+For the distal current $I_d(t)$ to be fully 
+`reconstructable' by the basal input, $x_d(t)$ has 
+to be a linear combination 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{align}
+x_d(t) &= \sum_{i=1}^N a_i x_{p,i}(t)
+\label{eq_x_d_a}
+\end{align}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+of the $x_{p,i}(t)$, where the $a_i$ are the
+components of a random vector $\mathbf{a}$ 
+of unit length.
+
+Given that we use with \eqref{eq_hebb_plast}
+a Hebbian learning scheme, one can expect that
+the direction and the magnitude of the principal 
+components of the basal input may affect the
+outcome of the simulation significantly:
+A large variance in the basal input 
 orthogonal to the `reconstruction vector' $\mathbf{a}$ 
-should act as a distraction for the plasticity and reduce the
-resulting alignment between $I_{\rm p}$ and $I_{\rm d}$. Therefore, we 
-applied a transformation to the input sequences $x_{p,i}(t)$
-that were parameterized by two quantities, a scaling factor $s$
-and the dimension $N_{dist}$ of a subspace of the basal input space.
-A set of $N_{dist}$ orthonormal basis vectors was randomly generated,
-which were also orthogonal to $\mathbf{a}$. 
-Withing this $N_{dist}$-dimensional subspace, the input sequences 
-$x_{p,i}(t)$ were then rescaled by the factor $s$.
+is a distraction for the plasticity. The
+observed temporal alignment between $I_p$ and 
+$I_d$ should hence suffer when such a 
+distraction is present. The situaton is illustrated 
+in Fig.~\ref{fig_illustration_classification}.
 
-After both $x_{p,i}(t)$ and $x_d(t)$ were generated,
-a simulation was run using all previously described plasticity
-mechanisms until the dynamic variables reached a stationary state.
-After this learning phase, another set of input sequences was generated
-using the previously described protocol and $\rho[I_{\rm p},I_{\rm d}]$ 
-was calculated. Note that plasticity was turned off in this phase. 
-This entire procedure allowed us to calculate $\rho[I_{\rm p},I_{\rm d}]$ 
-as a function of the distraction parameters $s$ and $N_{dist}$. 
-The results for both neuron models is shown in 
-Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
-basal inputs is $N=10$. One can observe a decorrelation transition for 
+In order to test the effects of distracting directions,
+we applied a transformation to the input sequences $x_{p,i}(t)$.
+For the transformation two parameters are used,
+a scaling factor $s$ and the dimension $N_{\rm dist}$ 
+of the distracting subspace within the basal input 
+space. The $N_{\rm dist}$ randomly generated
+basis vectors are orthogonal to the superposition
+vector $\mathbf{a}$, as defined by \eqref{eq_x_d_a},
+and to each others. Within this $N_{\rm dist}$-dimensional 
+subspace, the input sequences $x_{p,i}(t)$ are
+rescaled subsequently by the factor $s$. 
+See Fig.~\ref{fig_illustration_classification}.
+After the learning phase, a second set of input 
+sequences $x_{p,i}(t)$ and $x_d(t)$ is generated for 
+testing purposes, using the identical protocol, and
+the cross correlation $\rho[I_p,I_d]$ 
+evaluated. During the testing phase plasticity is 
+turned off.
+
+The overall aim of our portocal is to evaluate
+the degree $\rho[I_p,I_d]$ to which
+the proximal current $I_p$ aligns in the
+temporal domain to the distal input $I_d$.
+We recall that this is a highly non-trivial
+question, given that the proximal synpatic
+weights are adapted via Hebbian plasticity, 
+see \eqref{eq_hebb_plast}. The error 
+$(I_p-I_d)^2$ does not
+enter the adaption rules employed.
+Results are presented in
+Fig.~\ref{fig_corr_dimension_scaling}
+as a function of the distraction parameters 
+$s$ and $N_{\rm dist}\in[0,N-1]$. The
+total number of basal inputs is $N=100$.
+
+For a comparison, in 
+Fig.~\ref{fig_corr_dimension_scaling}
+data for both the compartment model
+and for a point neuron are presented,
+as defined respectively by
+\eqref{eq_comp_model}
+and \eqref{eq_point_neuron}.
+One can observe a decorrelation transition for 
 both models. However, the compartment model supports a 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{corr_dimension_scaling_high_input_dim}
+\includegraphics[width=1.0\columnwidth]{corr_dimension_scaling_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input.} Color
-	encodes the Pearson correlation $\rho[I_p,I_d]$ for different
-	number of orthogonal distraction directions $N_{dist}$ 
-	and the corresponding scaling faction $s$.}
-\label{fig:corr_dimension_scaling}
+encoded the Pearson correlation $\rho[I_p,I_d]$ 
+between the proximal and distal input currents,
+$I_p$ and $I_d$. Data for a range of the orthogonal 
+distraction directions, $N_{\rm dist}\in[0,N-1]$, 
+and the scaling factor $s$, as defined in
+Fig.~\ref{fig_illustration_classification}.
+The overall number of basal inputs is $N=100$.
+The quantity $\Sigma_{\rho}$ depicted in the bar
+plot is the sum over all correlation values for the different tested values
+of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
+orange the point model.}
+\label{fig_corr_dimension_scaling}
 \end{figure}
@@
 For this purpose, we constructed presynaptic basal 
-input as illustrated in Fig.~\ref{fig:illustration_classification}.
+input as illustrated in Fig.~\ref{fig_illustration_classification}.
 Written in vector form, each sample from the basal 
@@
 \mathbf{x}_p(t) = \mathbf{b} + \mathbf{a}\left(c(t) + \sigma_a \zeta_a(t) \right) 
-+ s \cdot \sum_{i=1}^{N_{dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
++ s \cdot \sum_{i=1}^{N_{\rm dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
 \end{equation}
@@
 Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
-orthogonal basis of $N_{dist}$ unit vectors which---just as in 
+orthogonal basis of $N_{\rm dist}$ unit vectors which---just as in 
 Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
@@
 once before the beginning of a simulation run.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
-\centering
-\includegraphics[width=0.55\columnwidth]{illustration_classification}
-\caption{{\bf Input Space for a Simple Linear Classification Task.}
-	Two clusters of presynaptic basal activity were generated from 
-	multivariate Gaussian distributions. Here, $s$ denotes the standard
-	deviation orthogonal to the normal vector of the classification
-	hyperplane.}
-\label{fig:illustration_classification}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 For the classification task, we set up two output neurons receiving
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{classification_dimension_scaling_high_input_dim}
+\includegraphics[width=1.0\columnwidth]{classification_dimension_scaling_high_input_dim}
 \caption{{\bf Binary Classification Accuracy.}
 	Fraction of correctly classified patterns as illustrated in
-	Fig.~\ref{fig:illustration_classification}, see 
-	Section~\ref{sect:classification}.¸}
+	Fig.~\ref{fig_illustration_classification}, see 
+	Section~\ref{sect:classification}. The quantity $\Sigma_{\rm acc}$ depicted in the bar
+	plot is the sum over all accuracies for the different tested values
+	of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
+	orange the point model.}
 \label{fig:classification_accuracy}
@@
 
-The resulting accuracy as a function of $N_{dist}$ and $s$
+The resulting accuracy as a function of $N_{\rm dist}$ and $s$
 is shown in Fig.~\ref{fig:classification_accuracy}.
@@
 improvements can be observed for the point model
-(compare Fig.~\ref{fig:corr_dimension_scaling_bcm} 
-with Fig.~\ref{fig:corr_dimension_scaling}).
+(compare Fig.~\ref{fig_corr_dimension_scaling_bcm} 
+with Fig.~\ref{fig_corr_dimension_scaling}).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{classification_dimension_scaling_bcm_high_input_dim}
+\includegraphics[width=1.0\columnwidth]{classification_dimension_scaling_bcm_high_input_dim}
 \caption{{\bf Binary Classification Accuracy, BCM Rule.}
 	Fraction of correctly classified patterns as illustrated in
-	Fig.~\ref{fig:illustration_classification}, after training with
-	a BCM-like learning rule.¸}
+	Fig.~\ref{fig_illustration_classification}, after training with
+	a BCM-like learning rule. The quantity $\Sigma_{\rm acc}$ depicted in the bar
+	plot is the sum over all accuracies for the different tested values
+	of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
+	orange the point model.}
 \label{fig:classification_accuracy_bcm}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{corr_dimension_scaling_bcm_high_input_dim}
+\includegraphics[width=1.0\columnwidth]{corr_dimension_scaling_bcm_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input, BCM Rule.}
 Color encodes the Pearson correlation $\rho[I_p,I_d]$ for different
-	number of orthogonal distraction directions $N_{dist}$ 
+number of orthogonal distraction directions $N_{\rm dist}$ 
 and the corresponding scaling faction $s$ after training
-	with a BCM-like rule. Compare to Fig.~\ref{fig:corr_dimension_scaling}.}
-\label{fig:corr_dimension_scaling_bcm}
+with a BCM-like rule. The quantity $\Sigma_{\rho}$ depicted in the bar
+plot is the sum over all correlation values for the different tested values
+of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
+orange the point model. Compare to Fig.~\ref{fig_corr_dimension_scaling}.}
+\label{fig_corr_dimension_scaling_bcm}
 \end{figure}
@@
 the learning rule for the proximal weights in terms of an objective function.
-For this purpose, we further simplify \eqref{eq:comp_model} by replacing
+For this purpose, we further simplify \eqref{eq_comp_model} by replacing
 the sigmoid functions $\sigma(x)$ by a simple step function $\Theta(x)$. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
 	\eqref{eq:obj_func}. This corresponds to a combination of using 
-	\eqref{eq:comp_model} together with \eqref{eq:bcm_rule}. Note the ridge-like
+	\eqref{eq_comp_model} together with \eqref{eq:bcm_rule}. Note the ridge-like
 	structure along the $I_p$-$I_d$ diagonal, which supports the alignment between
@@
 has already been proposed in different variants
-\cite{Bengio2014,Lee2015,Guerguiev2017}. Yet, all of these approaches
+\citep{Bengio2014,Lee2015,Guerguiev2017}. Yet, all of these approaches
 assume a learning rule being dependent on an explicit error term 
@@
 $I_p$ and $I_d$ can be regarded as a form of canonical correlation
-analysis (CCA) \cite{Haerdle2007}. The idea of investigating CCA as a
+analysis (CCA) \citep{Haerdle2007}. The idea of investigating CCA as a
 possible mode of synaptic learning has previously been investigated by
@@
 input, even in the presence of distracting principal components.
+Apart from the advantage that this avoids the necessity of giving
+a biophysical interpretation of such cross-terms, it is also in line
+with the view that synaptic plasticity should be formulated
+in terms of local membrane voltage traces \citep{Clopath2010,Weissenberger2018}. 
+Distal compartments then only implicitly affect plasticity, 
+e.g.\ by facilitating spike initiation.
 
@@
 indicated, random top-down weights can be sufficient for
-successful credit assignment and learning \cite{Lillicrap2016,Guerguiev2017}.
+successful credit assignment and learning \citep{Lillicrap2016,Guerguiev2017}.
 Therefore, our results could potentially also be transferred to deeper
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{classification_correlation_dimension_scaling_high_input_dim}
+\includegraphics[width=1.0\columnwidth]{classification_correlation_dimension_scaling_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning.}
 	Correlation between proximal and distal Input after training as described in 
-	Sect.~\ref{sect:classification}.}
+	Sect.~\ref{sect:classification}. The quantity $\Sigma_{\rho}$ depicted in the bar
+	plot is the sum over all correlation values for the different tested values
+	of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
+	orange the point model.}
 \label{fig:classification_correlation}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.7\columnwidth]{classification_correlation_dimension_scaling_bcm_high_input_dim}
+\includegraphics[width=1.0\columnwidth]{classification_correlation_dimension_scaling_bcm_high_input_dim}
 \caption{{\bf Alignment between Basal and Apical Input after Binary Classification Learning
@@
 	training as described in Sect.~\ref{sect:classification}, but using the plasticity
-	rules given in Sect.~\ref{sect:non-hebbian}.}
+	rules given in Sect.~\ref{sect:non-hebbian}. The quantity $\Sigma_{\rho}$ depicted in the bar
+	plot is the sum over all correlation values for the different tested values
+	of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
+	orange the point model.}
 \label{fig:classification_correlation_bcm}
@@
 the German research foundation (DFG)
-\section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} 
- should be uploaded separately on submission, if there are Supplementary Figures, 
- please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
+%\section*{Supplemental Data}
+% \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} 
+% should be uploaded separately on submission, if there are Supplementary Figures, 
+% please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
 

### 2021-05-12 code/plotting/classification_correlation_dimension_scaling_bcm_high_input_dim.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set()
 from stdParams import *
-from src.plottools import gen_mesh_ax
+from src.plottools import gen_mesh_ax, colorbar
 plt.style.use('mpl_style.mplstyle')
 import os
 import sys
@@ -76,16 +76,23 @@ rho_mean = rho.mean(axis=3)
 dim_ax = gen_mesh_ax(dim)
 s_ax = gen_mesh_ax(s)
 
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.35))
+#fig, ax = plt.subplots(1,4, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
 
 pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
 pc1 = ax[1].pcolormesh(s_ax,dim_ax,rho_mean[:,:,1].T,rasterized=True,vmin=0.,vmax=1.)
+bars0 = ax[2].barh(dim-2.5,rho_mean[:,:,0].sum(axis=0),height=5,label="Comp.")
+bars1 = ax[2].barh(dim+2.5,rho_mean[:,:,1].sum(axis=0),height=5,label="Point")
+
+ax[2].legend()
 
 ax[0].set_xlabel(r'$s$')
 ax[1].set_xlabel(r'$s$')
 
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
+ax[0].set_ylabel(r'$N_{\rm dist}$')
+ax[1].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_xlabel(r'$\Sigma_{\rho}$')
 
 #ax[0].set_xlim(right=5.)
 #ax[1].set_xlim(right=5.)
@@ -93,9 +100,10 @@ ax[1].set_ylabel(r'$N_{dist}$')
 ax[0].set_title("Compartment Model",loc="right")
 ax[1].set_title("Point Model",loc="right")
 
[…]

### 2021-05-12 code/plotting/classification_correlation_dimension_scaling_high_input_dim.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set()
 from stdParams import *
-from src.plottools import gen_mesh_ax
+from src.plottools import gen_mesh_ax, colorbar
 plt.style.use('mpl_style.mplstyle')
 import os
 import sys
@@ -76,16 +76,22 @@ rho_mean = rho.mean(axis=3)
 dim_ax = gen_mesh_ax(dim)
 s_ax = gen_mesh_ax(s)
 
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.35))
 
 pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
 pc1 = ax[1].pcolormesh(s_ax,dim_ax,rho_mean[:,:,1].T,rasterized=True,vmin=0.,vmax=1.)
+bars0 = ax[2].barh(dim-2.5,rho_mean[:,:,0].sum(axis=0),height=5,label="Comp.")
+bars1 = ax[2].barh(dim+2.5,rho_mean[:,:,1].sum(axis=0),height=5,label="Point")
+
+ax[2].legend()
 
 ax[0].set_xlabel(r'$s$')
 ax[1].set_xlabel(r'$s$')
 
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
+ax[0].set_ylabel(r'$N_{\rm dist}$')
+ax[1].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_xlabel(r'$\Sigma_{\rho}$')
 
 #ax[0].set_xlim(right=5.)
 #ax[1].set_xlim(right=5.)
@@ -93,9 +99,9 @@ ax[1].set_ylabel(r'$N_{dist}$')
 ax[0].set_title("Compartment Model",loc="right")
 ax[1].set_title("Point Model",loc="right")
 
-plt.colorbar(pc0,cax=ax[2])
+colorbar(pc1)
 
-fig.tight_layout(pad=0.2)
[…]

### 2021-05-12 code/plotting/classification_dimension_scaling_bcm_high_input_dim.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set()
 from stdParams import *
-from src.plottools import gen_mesh_ax
+from src.plottools import gen_mesh_ax, colorbar
 plt.style.use('mpl_style.mplstyle')
 import os
 import sys
@@ -58,16 +58,22 @@ n_sweep = perf.shape[2]
 
 perf_mean = perf.mean(axis=3)
 
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.35))
 
 pc0 = ax[0].pcolormesh(s_ax,dim_ax,perf_mean[:,:,0].T,rasterized=True,vmin=0.5,vmax=1.)
 pc1 = ax[1].pcolormesh(s_ax,dim_ax,perf_mean[:,:,1].T,rasterized=True,vmin=0.5,vmax=1.)
+bars0 = ax[2].barh(dim-2.5,perf_mean[:,:,0].sum(axis=0),height=5,label="Comp.")
+bars1 = ax[2].barh(dim+2.5,perf_mean[:,:,1].sum(axis=0),height=5,label="Point")
+
+ax[2].legend()
 
 ax[0].set_xlabel(r'$s$')
 ax[1].set_xlabel(r'$s$')
 
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
+ax[0].set_ylabel(r'$N_{\rm dist}$')
+ax[1].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_xlabel(r'$\Sigma_{\rm acc}$')
 
 #ax[0].set_xlim(right=5.)
 #ax[1].set_xlim(right=5.)
@@ -75,9 +81,9 @@ ax[1].set_ylabel(r'$N_{dist}$')
 ax[0].set_title("Compartment Model",loc="right")
 ax[1].set_title("Point Model",loc="right")
 
-plt.colorbar(pc0,cax=ax[2])
+colorbar(pc1)
 
-fig.tight_layout(pad=0.2)
[…]

### 2021-05-12 code/plotting/classification_dimension_scaling_high_input_dim.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set()
 from stdParams import *
-from src.plottools import gen_mesh_ax
+from src.plottools import gen_mesh_ax, colorbar
 plt.style.use('mpl_style.mplstyle')
 import os
 import sys
@@ -58,16 +58,22 @@ n_sweep = perf.shape[2]
 
 perf_mean = perf.mean(axis=3)
 
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.35))
 
 pc0 = ax[0].pcolormesh(s_ax,dim_ax,perf_mean[:,:,0].T,rasterized=True,vmin=0.5,vmax=1.)
 pc1 = ax[1].pcolormesh(s_ax,dim_ax,perf_mean[:,:,1].T,rasterized=True,vmin=0.5,vmax=1.)
+bars0 = ax[2].barh(dim-2.5,perf_mean[:,:,0].sum(axis=0),height=5,label="Comp.")
+bars1 = ax[2].barh(dim+2.5,perf_mean[:,:,1].sum(axis=0),height=5,label="Point")
+
+ax[2].legend()
 
 ax[0].set_xlabel(r'$s$')
 ax[1].set_xlabel(r'$s$')
 
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
+ax[0].set_ylabel(r'$N_{\rm dist}$')
+ax[1].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_xlabel(r'$\Sigma_{\rm acc}$')
 
 #ax[0].set_xlim([0.1,5.])
 #ax[1].set_xlim([0.1,5.])
@@ -75,9 +81,9 @@ ax[1].set_ylabel(r'$N_{dist}$')
 ax[0].set_title("Compartment Model",loc="right")
 ax[1].set_title("Point Model",loc="right")
 
-plt.colorbar(pc1,cax=ax[2])
+colorbar(pc1)
 
-fig.tight_layout(pad=0.2)
[…]

### 2021-05-12 code/plotting/correlation_dimension_scaling_bcm_high_input_dim.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set()
 from stdParams import *
-from src.plottools import gen_mesh_ax
+from src.plottools import gen_mesh_ax, colorbar
 plt.style.use('mpl_style.mplstyle')
 import os
 
@@ -60,16 +60,22 @@ n_sweep = rho.shape[2]
 
 rho_mean = rho.mean(axis=3)
 
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.35))
 
 pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
 pc1 = ax[1].pcolormesh(s_ax,dim_ax,rho_mean[:,:,1].T,rasterized=True,vmin=0.,vmax=1.)
+bars0 = ax[2].barh(dim-2.5,rho_mean[:,:,0].sum(axis=0),height=5,label="Comp.")
+bars1 = ax[2].barh(dim+2.5,rho_mean[:,:,1].sum(axis=0),height=5,label="Point")
+
+ax[2].legend()
 
 ax[0].set_xlabel(r'$s$')
 ax[1].set_xlabel(r'$s$')
 
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
+ax[0].set_ylabel(r'$N_{\rm dist}$')
+ax[1].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_xlabel(r'$\Sigma_{\rho}$')
 
 ax[0].set_xlim(right=4.)
 ax[1].set_xlim(right=4.)
@@ -77,9 +83,9 @@ ax[1].set_xlim(right=4.)
 ax[0].set_title("Compartment Model",loc="right")
 ax[1].set_title("Point Model",loc="right")
 
-plt.colorbar(pc1,cax=ax[2])
+colorbar(pc1)
 
-fig.tight_layout(pad=0.2)
[…]

### 2021-05-12 code/plotting/correlation_dimension_scaling_high_input_dim.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 sns.set()
 from stdParams import *
-from src.plottools import gen_mesh_ax
+from src.plottools import gen_mesh_ax, colorbar
 plt.style.use('mpl_style.mplstyle')
 import os
 import sys
@@ -58,16 +58,22 @@ n_sweep = rho.shape[2]
 
 rho_mean = rho.mean(axis=3)
 
-fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [15,15,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.45))
+fig, ax = plt.subplots(1,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.35))
 
 pc0 = ax[0].pcolormesh(s_ax,dim_ax,rho_mean[:,:,0].T,rasterized=True,vmin=0.,vmax=1.)
 pc1 = ax[1].pcolormesh(s_ax,dim_ax,rho_mean[:,:,1].T,rasterized=True,vmin=0.,vmax=1.)
+bars0 = ax[2].barh(dim-2.5,rho_mean[:,:,0].sum(axis=0),height=5,label="Comp.")
+bars1 = ax[2].barh(dim+2.5,rho_mean[:,:,1].sum(axis=0),height=5,label="Point")
+
+ax[2].legend()
 
 ax[0].set_xlabel(r'$s$')
 ax[1].set_xlabel(r'$s$')
 
-ax[0].set_ylabel(r'$N_{dist}$')
-ax[1].set_ylabel(r'$N_{dist}$')
+ax[0].set_ylabel(r'$N_{\rm dist}$')
+ax[1].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_ylabel(r'$N_{\rm dist}$')
+ax[2].set_xlabel(r'$\Sigma_{\rho}$')
 
 ax[0].set_xlim(right=4.)
 ax[1].set_xlim(right=4.)
@@ -75,9 +81,9 @@ ax[1].set_xlim(right=4.)
 ax[0].set_title("Compartment Model",loc="right")
 ax[1].set_title("Point Model",loc="right")
 
-plt.colorbar(pc1,cax=ax[2])
+colorbar(pc1)
 
-fig.tight_layout(pad=0.2)
[…]

### 2021-05-12 code/simulation/classification_dimension_scaling_bcm_high_input_dim.py

@@ -14,9 +14,6 @@ N_sweep_distraction_dimension = 9
 N_samples = 1
 distract_scaling = np.linspace(0.,10.,N_sweep_distraction_scaling)
 
-
-
-
 N_p = 100
 N_out = 2
 

### 2021-05-12 code/src/plottools.py

@@ -5,3 +5,15 @@ def gen_mesh_ax(x):
     x_mesh = np.append(x,2.*x[-1] - x[-2])
     x_mesh -= .5 * (x_mesh[1] - x_mesh[0])
     return x_mesh
+    
+def colorbar(mappable):
+    from mpl_toolkits.axes_grid1 import make_axes_locatable
+    import matplotlib.pyplot as plt
+    last_axes = plt.gca()
+    ax = mappable.axes
+    fig = ax.figure
+    divider = make_axes_locatable(ax)
+    cax = divider.append_axes("right", size="5%", pad=0.05)
+    cbar = fig.colorbar(mappable, cax=cax)
+    plt.sca(last_axes)
+    return cbar

### 2021-05-25 notes/frontiers_coincidence_learning/frontiers.tex
.

@@
 %---------------------------------------%
-\subsection{Alignment between Basal and Apical Inputs}
+\subsection{Unsupervised Alignment between Basal and Apical Inputs}
 \label{sect:alignment}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{corr_dimension_scaling_high_input_dim}
+\caption{{\bf Unsupervised Alignment between 
+Basal and Apical Input.} Color
+encoded the Pearson correlation $\rho[I_p,I_d]$ 
+between the proximal and distal input currents,
+$I_p$ and $I_d$. Data for a range of the orthogonal 
+distraction directions, $N_{\rm dist}\in[0,N-1]$, 
+and the scaling factor $s$, as defined in
+Fig.~\ref{fig_illustration_classification}.
+The overall number of basal inputs is $N=100$.
+The quantity $\Sigma_{\rho}$ depicted in the bar
+plot is the sum, for a given $N_{\rm dist}$,
+over the correlation values $s=0,\,0.5,\,1.0\,..$ 
+used for the simulation run. Blue bars represents 
+the compartment model, orange the point model.}
+%\Delta s = 10/(N_s-1) von 0 bis 10.
+% > D.h.  s\in[0,10],   N_s = 20 ?
+\label{fig_corr_dimension_scaling}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 The overall aim of our portocal is to evaluate
@@
 and for a point neuron are presented,
-as defined respectively by
-\eqref{eq_comp_model}
-and \eqref{eq_point_neuron}.
-One can observe a decorrelation transition for 
-both models. However, the compartment model supports a 
-significantly stronger distraction in terms of the scaling 
-factor $s$ as compared to the point model. This is a first confirmation 
-of our hypothesis that nonlinear interactions between basal and 
-apical input could improve learning guided by top-down signals.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{corr_dimension_scaling_high_input_dim}
-\caption{{\bf Alignment between Basal and Apical Input.} Color
-encoded the Pearson correlation $\rho[I_p,I_d]$ 
-between the proximal and distal input currents,
-$I_p$ and $I_d$. Data for a range of the orthogonal 
-distraction directions, $N_{\rm dist}\in[0,N-1]$, 
-and the scaling factor $s$, as defined in
-Fig.~\ref{fig_illustration_classification}.
-The overall number of basal inputs is $N=100$.
-The quantity $\Sigma_{\rho}$ depicted in the bar
-plot is the sum over all correlation values for the different tested values
-of $s$ under a given $N_{\rm dist}$. Blue bar represents the compartment model,
-orange the point model.}
-\label{fig_corr_dimension_scaling}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+as defined respectively by \eqref{eq_comp_model}
+and \eqref{eq_point_neuron}. A decorrelation 
+transition as a function of the distraction 
+scaling paramerer $s$ is observed for both 
+models. However, the compartment model is able
+to handle a significantly stronger distraction 
+as compared to the point model. These findings
+support the hypothesis examined here, namely
+that nonlinear interactions between basal and 
+apical input improve learning guided by top-down signals.
 
@@
 For this purpose, we constructed presynaptic basal 
-input as illustrated in Fig.~\ref{fig_illustration_classification}.
+input $x_p(t)$ as illustrated in 
+Fig.~\ref{fig_illustration_classification}.
 Written in vector form, each sample from the basal 
@@
 %---------------------------------------%
-\subsection{Non-Hebbian Learning Rules}
+\subsection{Generalized Hebbian Learning Rules}
 \label{sect:non-hebbian}
@@
 
-Instead of Hebbian learning, we also considered a BCM-like
-learning rule for the basal weights \citep{Bienenstock1982,Intrator1992}.
-The form of the BCM-rule we consider here reads
+Instead of basic Hebbian learning, we also considered 
+a BCM-like learning rule for the basal weights 
+\citep{Bienenstock1982,Intrator1992}.
+The form of the BCM-rule used here reads
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-25 code/plotting/classification_correlation_dimension_scaling_high_input_dim_composite.py

@@ -1,0 +1,130 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax, colorbar
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+plt.rc('text.latex', preamble=r'\renewcommand{\familydefault}{\sfdefault}')
+
+folders = ["classification_dimension_scaling_high_input_dim"
+,"classification_dimension_scaling_bcm_high_input_dim"]
+
+fig, ax = plt.subplots(2,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.65))
+
+for i in range(2):
+
+    files = os.listdir(os.path.join(DATA_DIR,folders[i]))
+
+    data = []
+
+    for file in files:
+        data.append(np.load(os.path.join(DATA_DIR,folders[i]+"/"+file)))
+
+    perf = []
+    Ip = []
+    labels = []
+    dim = []
+    s = []
+
+    for dat in data:
+        perf.append(dat["perf"])
+        Ip.append(dat["I_p_samples"])
+        labels.append(dat["labels_samples"])
+        dim.append(dat["distract_dimension"])
+        s.append(dat["distract_scaling"])
+
+    for k in range(1,len(perf)):
+        if(not(np.array_equal(Ip[k].shape,Ip[k-1].shape))):
+            print("Ip arrays do not match!")
+            sys.exit()
+        if(not(np.array_equal(labels[k].shape,labels[k-1].shape))):
+            print("labels arrays do not match!")
+            sys.exit()
+        if(not(np.array_equal(dim[k],dim[k-1]))):
+            print("dim arrays do not match!")
+            sys.exit()
+        if(not(np.array_equal(s[k],s[k-1]))):
+            print("scaling arrays do not match!")
+            sys.exit()
+
+    perf = np.array(perf)
+    Ip = np.array(Ip)
+    labels = np.array(labels)
+    dim = dim[0]
+    s = s[0]
+
[…]

### 2021-05-25 code/plotting/correlation_dimension_scaling_high_input_dim_composite.py

@@ -1,0 +1,111 @@
+#! /usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set()
+from stdParams import *
+from src.plottools import gen_mesh_ax, colorbar
+plt.style.use('mpl_style.mplstyle')
+import os
+import sys
+
+plt.rc('text.latex', preamble=r'\renewcommand{\familydefault}{\sfdefault}')
+
+folders = ["correlation_dimension_scaling_high_input_dim"
+,"correlation_dimension_scaling_bcm_high_input_dim"]
+
+fig, ax = plt.subplots(2,3,gridspec_kw={'width_ratios': [1,1,1]}, figsize=(FIG_WIDTH,FIG_WIDTH*0.65))
+
+for i in range(2):
+
+    files = os.listdir(os.path.join(DATA_DIR,folders[i]))
+
+    data = []
+
+    for file in files:
+        data.append(np.load(os.path.join(DATA_DIR,folders[i]+"/"+file)))
+
+    rho = []
+    dim = []
+    s = []
+
+    for dat in data:
+        rho.append(dat["align_corrcoef"])
+        dim.append(dat["distract_dimension"])
+        s.append(dat["distract_scaling"])
+
+    for k in range(1,len(rho)):
+        if(not(np.array_equal(rho[k].shape,rho[k-1].shape))):
+            print("corr. arrays do not match!")
+            sys.exit()
+        if(not(np.array_equal(dim[k],dim[k-1]))):
+            print("dim arrays do not match!")
+            sys.exit()
+        if(not(np.array_equal(s[k],s[k-1]))):
+            print("scaling arrays do not match!")
+            sys.exit()
+
+    rho = np.array(rho)
+    dim = dim[0]
+    s = s[0]
+
+    n_total_samples = rho.shape[0]*rho.shape[3]
+
+    rho_flatten = np.ndarray((rho.shape[1],rho.shape[2],rho.shape[4],n_total_samples))
+
+    for k in range(rho.shape[0]):
+        for l in range(rho.shape[3]):
+            rho_flatten[:,:,:,k*rho.shape[3]+l] = rho[k,:,:,l,:]
+
[…]
